---
title: "Analyse terrestrial indicators"
author: "Simone Stevenson"
date: "20/09/2021"
output:
  html_document: default
  pdf_document: default
---

#' TODO: Fix axis labels + titles on bootstrapped plots
#' TODO: Add panel labels
#' TODO: Change blue and red to other colours
#' TODO: Add vector art for scenarios (lion, wildebeest, plant)

```{r setup, include=FALSE}

# https://bookdown.org/yihui/rmarkdown/r-code.html

knitr::opts_chunk$set(echo = FALSE, message = FALSE,
  warnings = FALSE, error = FALSE)

```

Directory path to git repo

Simone's deakin laptop

cd "C:\\Users\\ssteven\\OneDrive - Deakin University\\Deakin\\Chapter_3_indicator_testing\\madingley_terrestrial_indicators"

Data structure

Input data is structured hierarchically as follows:
- Location (1)
-- Scenarios (3)
--- Replicates (25)

Output indicator data is calculated at the same level, so each indicator should
have values for:
- Location (1)
-- Scenarios (3)
--- Replicates (25)

# Possible useful ref?
https://www.frontiersin.org/articles/10.3389/fevo.2018.00149/full

```{r libraries}

## Data wrangling
library(tidyverse)
library(tidylog)
library(zoo)

## Analysis
library(mgcv)
library(nlme)
library(MASS)
library(datawizard)

## Viz
library(ggpubr)
library(cowplot)
library(gridExtra)
library(gratia)
library(kableExtra)

```

```{r template chunk}

```

```{r ggplot theme}

boot_theme <- theme(panel.background = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              legend.position = "bottom",
              legend.key.size = unit(0.25, 'cm'),
              plot.title = element_text(size = 10, face = "bold"),
              legend.title=element_text(size=6), 
              legend.text=element_text(size=5),
              axis.line = element_line(colour = "black"),
              axis.title.x = element_text(size=8),
              axis.title.y = element_text(size=8),
              axis.text.x = element_text(size=8),
              axis.text.y = element_text(size=8),
              axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())

```

```{r get test gam data }

#' @param model - gam object
#' @param data - the input pressure response time step data for the model
#' @return a dataframe containing predicted, observed, first and second derivatives

get_test_gam_data <- function(data, model) {
  
# Get observed indicator values

observed <- data$indicator_score

# Make a dataframe of predictor vals (ie timesteps)

pdat1 <- data.frame(annual_time_step = seq(1,290,1))

# Get predicted values for each time step
# p2 <- predict.gam(model, pdat1, 
#                   se.fit = TRUE, response = TRUE)

p2 <- predict(model, pdat1, 
                  se.fit = TRUE, response = TRUE)

p3 <- predict.gam(model, pdat1, 
                  se.fit = TRUE)

# # Get the deviance residuals
deviance_residuals <- residuals(model, type = "working")

# fv <- predict(model,type="terms") ## get term estimates
#      
# ## compute partial residuals for first smooth...          
# deviance_residuals <- residuals(model,type="working", response = TRUE) + fv[,1]

#Bind them together with observed vals

pdat <- cbind(pdat1, predicted = p2$fit, predicted_nr = p3$fit, se2 = p2$se.fit, observed, 
              deviance_residuals)

#Get upper and lower bounds for predicted values (replace with bootstraps for selected model)

df.res <- df.residual(model)
crit.t <- qt(0.025, df.res, lower.tail = FALSE)
pdat <- transform(pdat,
                  upper_pred = predicted_nr + (crit.t * se2),
                  lower_pred = predicted_nr - (crit.t * se2))

# Calculate the derivatives and add that too
derivs_1 <- as.data.frame(derivatives(model, order = 1, 
                   n = 290, interval = "confidence") ) %>% 
                   dplyr::select(derivative, lower, upper) %>% 
                   rename(first_derivative = derivative,
                          first_lower = lower,
                          first_upper = upper)

# Calculate the derivatives and add that too
derivs_2 <- as.data.frame(derivatives(model, order = 2, 
                   n = 290, interval = "confidence") ) %>% 
                   dplyr::select(derivative, lower, upper) %>% 
                   rename(second_derivative = derivative,
                          second_lower = lower,
                          second_upper = upper)
# Bind it all together
temp <- cbind(pdat, derivs_1, derivs_2)

# Get the indicator we are modeling
indicator <- data$indicator[1]
scenario <- data$scenario[1]

# Make the final plot data
test_gam_data <- temp %>% 
                 mutate(first_significant = ifelse(first_lower <= 0 & 
                                                  first_upper >= 0, FALSE, TRUE),
                        positive_trend = ifelse(first_significant == TRUE & 
                                                 first_derivative > 0,
                                            predicted_nr,NA),
                        negative_trend = ifelse(first_significant == TRUE & 
                                                 first_derivative < 0,
                                            predicted_nr, NA),
                        positive_der_1 = ifelse(first_significant == TRUE & 
                                                  first_derivative > 0,
                                            first_derivative,NA),
                        negative_der_1 = ifelse(first_significant == TRUE & 
                                                  first_derivative < 0,
                                            first_derivative, NA),
                        second_significant = ifelse(second_lower <= 0 & 
                                                  first_upper >= 0, FALSE, TRUE),
                        positive_tp = ifelse(second_significant == TRUE & 
                                             second_derivative > 0,
                                            predicted_nr,NA),
                        negative_tp = ifelse(second_significant == TRUE & 
                                            second_derivative < 0,
                                            predicted_nr, NA),
                        positive_der_2 = ifelse(second_significant == TRUE & 
                                                second_derivative > 0,
                                            second_derivative,NA),
                        negative_der_2 = ifelse(second_significant == TRUE & 
                                                second_derivative < 0,
                                            first_derivative, NA),
                        indicator = indicator,
                        scenario = scenario,
                        kvalue = model$rank,
                        r2 = summary.gam(model)$r.sq) 

}
```

```{r plot test gam}

plot_test_gam <- function(test_gam_data) {

if(all(is.na(test_gam_data$positive_trend))) {

gam_plot <- ggplot(data = test_gam_data) +
            #geom_point(aes(x = annual_time_step, y = deviance_residuals), alpha = 0.2) +
            geom_point(aes(x = annual_time_step, y = observed), alpha = 0.2) +
            geom_rug(sides="b", aes(x = annual_time_step)) +
            # Plot the smoothed indicator line
            geom_line(aes(x = annual_time_step, y = predicted_nr)) +
            # Add the confidence intervals
            geom_ribbon(aes(x = annual_time_step, ymin = lower_pred, ymax = upper_pred),
                        alpha = 0.3) +
            # Add the significant trends based on first derivative
            #geom_line(aes(x = annual_time_step, y = positive_trend), col = "blue") +
            geom_line(aes(x = annual_time_step, y = negative_trend), col = "red") +
            labs(y = test_gam_data$indicator[1],
                 title = paste(test_gam_data$scenario[1], 
                               test_gam_data$indicator[1], "k =", 
                               test_gam_data$kvalue[1], sep = " ")) +
            theme_classic()
  
} else {
  
gam_plot <- ggplot(data = test_gam_data) +
            #geom_point(aes(x = annual_time_step, y = deviance_residuals), alpha = 0.2) +
            geom_point(aes(x = annual_time_step, y = observed), alpha = 0.2) +
            geom_rug(sides="b", aes(x = annual_time_step)) +
            # Plot the smoothed indicator line
            geom_line(aes(x = annual_time_step, y = predicted_nr)) +
            # Add the confidence intervals
            geom_ribbon(aes(x = annual_time_step, ymin = lower_pred, ymax = upper_pred),
                        alpha = 0.3) +
            # Add the significant trends based on first derivative
            geom_line(aes(x = annual_time_step, y = positive_trend), col = "blue") +
            geom_line(aes(x = annual_time_step, y = negative_trend), col = "red") +
            labs(y = test_gam_data$indicator[1],
                 title = paste(test_gam_data$scenario[1], 
                               test_gam_data$indicator[1], "k =", 
                               test_gam_data$kvalue[1], sep = " ")) +
            theme_classic()

return(gam_plot)

  }
}


```

```{r compare trends}
# Loop through each model and identify which time steps show a significant 
# positive or negative trend

compare_trends <- function(list_of_model_data) {

trend_data <- list()

for (j in seq_along(list_of_model_data)) {
  
trend_data[[j]] <- list_of_model_data[[j]] %>% 
                   dplyr::select(annual_time_step, positive_trend, 
                                 negative_trend) %>% 
                   mutate(k_value = kvals[[j]],
                          direction = ifelse(!is.na(positive_trend),
                                             "positive",
                                            ifelse(!is.na(negative_trend),
                                             "negative", NA)))

}

# Convert  into a single data frame

trend_df_all <- do.call(rbind, trend_data)

# Check which time steps show a significant trend regardless of model
# k value (ie are the robust significant changes at the expected time points?)

direction_df <- trend_df_all %>% 
                dplyr::select(annual_time_step, k_value, direction) %>% 
                group_by(annual_time_step) %>% 
                summarise(positive_across_all_models = sum(direction == "positive"),
                          negative_across_all_models = sum(direction == "negative")) %>% 
                filter(positive_across_all_models == 4|
                       negative_across_all_models == 4) %>% 
                mutate(positive_across_all_models = ifelse(positive_across_all_models == 4, 1, 0),
                       negative_across_all_models = ifelse(negative_across_all_models == 4, 1, 0))

return(direction_df)

}

```

```{r bootstrap selected model}

bootstrap_selected_model <- function(data, model, model_data, numboots) {
  
bootstrap_mods <- list()
bootstrap_mod_data <- list()
replicate <- list()

for (k in 1:numboots) {
        
# Take k number of random samples from the input data
replicate[[k]] <- slice_sample(data, 
                          prop = 1, replace = TRUE) 
        

# Fit the model with the data
bootstrap_mods[[k]] <- gam(indicator_score ~ s(annual_time_step, k = model$rank),
          select = TRUE,
          method = "GCV.Cp",
          data = replicate[[k]])

# Get the predicted values etc

obs <- replicate[[k]]$indicator_score

# Make a dataframe of predictor vals (ie timesteps)

pdat1 <- data.frame(annual_time_step = seq(1,290,1))

# Get predicted values for each time step
p2 <- predict.gam(bootstrap_mods[[k]], pdat1,
                  se.fit = TRUE, response = TRUE)

# Get the deviance residuals
deviance_residuals <- residuals(bootstrap_mods[[k]], type = "deviance", 
                                response = TRUE)

#Bind them together with observed vals

pdat <- cbind(pdat1, predicted = p2$fit, se2 = p2$se.fit, obs, 
              deviance_residuals) %>%  
                  mutate(replicate = k)

# Calculate the derivatives and add that too
gratia_derivs_1 <- as.data.frame(derivatives(bootstrap_mods[[k]], order = 1, 
                   n = 290, interval = "confidence") ) %>% 
                   dplyr::select(derivative, lower, upper) %>% 
                   rename(first_derivative = derivative,
                          first_lower = lower,
                          first_upper = upper)

# Calculate the derivatives and add that too
gratia_derivs_2 <- as.data.frame(derivatives(bootstrap_mods[[k]], order = 2, 
                   n = 290, interval = "confidence") ) %>% 
                   dplyr::select(derivative, lower, upper) %>% 
                   rename(second_derivative = derivative,
                          second_lower = lower,
                          second_upper = upper)

# Bind it all together
bootstrap_mod_data[[k]] <- cbind(pdat, gratia_derivs_1, gratia_derivs_2)

# print(k)

}

ci_df <- do.call(rbind, bootstrap_mod_data) %>% 
             #Group by timestep
             group_by(annual_time_step) %>% 
             arrange(predicted) %>% 
             summarise(predicted_lower = quantile(predicted, 
                                      probs = 0.025),
                       predicted_upper = quantile(predicted, 
                                      probs = 0.975)) 

first_deriv_df <- do.call(rbind, bootstrap_mod_data) %>% 
             #Group by timestep
             group_by(annual_time_step) %>% 
             # Arrange first deriv values from each bootstrap in order
             arrange(first_derivative) %>% 
             # Set bottom 25th and top 75th quantiles as the lower and upper CIs of the first derivative
             summarise(first_lower = quantile(first_derivative, 
                                      probs = 0.025),
                       first_upper = quantile(first_derivative, 
                                      probs = 0.975)) %>% 
             # Add in the original first derivative from the full dataset
             merge(model_data[c("annual_time_step", "predicted",
                                "first_derivative")],
                   by = "annual_time_step") %>% 
             # Check if the derivative CIs include zero (if so trend is not significant)
             mutate(significant = ifelse(first_lower < 0 & first_upper > 0, 
                                         FALSE, TRUE),
             # If trend is significant and first deriv > 0, get the the predicted indicator 
             # trend values that are positive
                    positive_trend = ifelse(significant == TRUE & first_derivative > 0,
                                            predicted,NA),
             # If trend is significant and first deriv > 0, get the the predicted indicator 
             # trend values that are negative
                    negative_trend = ifelse(significant == TRUE & first_derivative < 0,
                                            predicted, NA),
             # If trend is significant and first deriv > 0, get the the predicted first 
             # derivative values that are positive
                    positive_der1 = ifelse(significant == TRUE & first_derivative > 0,
                                            first_derivative,NA),
             # If trend is significant and first deriv > 0, get the the predicted first 
             # derivative values that are negative
                    negative_der1 = ifelse(significant == TRUE & first_derivative < 0,
                                            first_derivative, NA)) %>% 
            dplyr::select(-predicted)

second_deriv_df <- do.call(rbind, bootstrap_mod_data) %>% 
             #Group by timestep
             group_by(annual_time_step) %>% 
             # Order second derivative values 
             arrange(second_derivative) %>% 
             summarise(second_lower = quantile(second_derivative, 
                                      probs = 0.025),
                       second_upper = quantile(second_derivative, 
                                      probs = 0.975))%>% 
             # Add in the original first derivative from the full dataset
             merge(model_data[c("annual_time_step", "predicted",
                                "second_derivative")],
                   by = "annual_time_step") %>% 
             mutate(significant_2 = ifelse(second_lower < 0 & second_upper > 0, 
                                         FALSE, TRUE),
                    positive_tp = ifelse(significant_2 == TRUE & second_derivative > 0,
                                            predicted,NA),
                    negative_tp = ifelse(significant_2 == TRUE & second_derivative < 0,
                                            predicted, NA),
                    positive_der2 = ifelse(significant_2 == TRUE & second_derivative > 0,
                                            second_derivative,NA),
                    negative_der2 = ifelse(significant_2 == TRUE & second_derivative < 0,
                                            second_derivative, NA)) %>% 
            dplyr::select(-predicted)


new <- model_data %>% 
     dplyr::select(annual_time_step, observed, predicted, deviance_residuals, 
                   indicator, scenario, kvalue) %>% 
     merge(ci_df, by ="annual_time_step") %>% 
     merge(first_deriv_df, by ="annual_time_step") %>% 
     merge(second_deriv_df, by ="annual_time_step") #%>% 
    # merge(model_data[c("annual_time_step", "kvalue")], by = "annual_time_step")

}


```

```{r plot bootstrapped model}

plot_bootstrapped_model <- function(data, ylimits){
  
indicator <- data$indicator[1]

boot_base_plot <- ggplot(data = data) +
                  geom_rect(aes(xmin = 95, xmax = 105,
                                ymin = -Inf, ymax = Inf), fill = "grey70", 
                            alpha = 0.4) +
                  geom_rect(aes(xmin = 195, xmax = 205,
                                ymin = -Inf, ymax = Inf), fill = "grey70", 
                            alpha = 0.4)+
                  geom_point(aes(x = annual_time_step, y = observed), 
                                       alpha = 0.2, size = 1) +
                  geom_line(aes(x = annual_time_step, y = observed), 
                            alpha = 0.2) +
                  geom_line(aes(x = annual_time_step, y = predicted)) +
                  geom_ribbon(aes(x = annual_time_step, ymin = predicted_lower, 
                                            ymax = predicted_upper),
                                        alpha = 0.3) +
                  labs(x = "Time (years)",
                        y = indicator) + 
                  geom_vline(xintercept=c(100,200), linetype = "dotted") +
                  coord_cartesian(ylim = ylimits)
                # + annotate("text", x = 295, y = 1.1, 
                #            label = paste("Dev. explained =",
                #                          dev_explained,sep = " "), 
                #            size = 4)

if(all(is.na(data$positive_trend))) {

boot_plot <-  boot_base_plot +
              geom_line(aes(x = annual_time_step, y = negative_trend), 
                        col = "#440154FF") +
              boot_theme 
  
} else if(all(is.na(data$robust_positive_trend))) {

boot_plot <- boot_base_plot +
   geom_line(aes(x = annual_time_step, y = positive_trend), col = "#2A788EFF") +
  geom_line(aes(x = annual_time_step, y = negative_trend), col = "#440154FF") +
  geom_line(aes(x = annual_time_step, y = robust_negative_trend), 
            col = "#440154FF", size = 2) +
  boot_theme  
 
  
} else if (all(is.na(data$robust_negative_trend))) {

boot_plot <- boot_base_plot +
   geom_line(aes(x = annual_time_step, y = positive_trend), col = "#2A788EFF") +
  geom_line(aes(x = annual_time_step, y = negative_trend), col = "#440154FF") +
  geom_line(aes(x = annual_time_step, y = robust_positive_trend), 
            col = "#2A788EFF", size = 2) +
  boot_theme 
 
} else {

  boot_plot <- boot_base_plot +
  geom_line(aes(x = annual_time_step, y = positive_trend), col = "#2A788EFF") +
  geom_line(aes(x = annual_time_step, y = negative_trend), col = "#440154FF") +
  geom_line(aes(x = annual_time_step, y = robust_positive_trend), 
            col = "#2A788EFF", size = 2) +
  geom_line(aes(x = annual_time_step, y = robust_negative_trend), 
            col = "#440154FF", size = 2) +
  boot_theme 
  
  return(boot_plot)
  
  }
}

```

# Inputs

```{r data, results=FALSE }

# Set up paths ----

location <- 'Serengeti'

indicators_project <- "N:/Quantitative-Ecology/Indicators-Project"

# Analysis inputs are the outputs of 2_calculate_madingley_indicators.R

analysis_inputs <- "N:/Quantitative-Ecology/Indicators-Project/Serengeti/Outputs_from_indicator_code/Indicator_outputs/general/"

# Inputs ----

# Specify the date of inputs you want to use

input_date <- "2021-09-30"

# Specify how many bootstraps you want to use

boots <- 1

# Get date to label outputs

output_date <- Sys.Date() 

scenarios_title <- list("Baseline", "Land use", 
                  "Carnivore harvesting", 
                  "Herbivore harvesting")

disturbance_string <- c("pre-disturbance", "disturbance", "post-disturbance")

disturbance_factor <- factor(disturbance_string, ordered = TRUE,
                              levels = c("pre-disturbance", 
                                         "disturbance", 
                                         "post-disturbance"))

# Set up output folders

## Include a folder to store the version of inputs used to get the outputs

analysis_inputs_folder <- file.path(indicators_project,  
                           "/Serengeti/Outputs_from_analysis_code/Analysis_inputs",
                           output_date)

if( !dir.exists( file.path(analysis_inputs_folder) ) ) {
  dir.create( file.path(analysis_inputs_folder), recursive = TRUE )
  
}

analysis_outputs_folder <- file.path(indicators_project, 
                                     "/Serengeti/Outputs_from_analysis_code/Analysis_outputs",
                                     output_date)

if( !dir.exists( file.path(analysis_outputs_folder) ) ) {
  dir.create( file.path(analysis_outputs_folder), recursive = TRUE )
  
}

manuscript_plots_folder <- file.path(indicators_project, 
                                     "/Serengeti/Outputs_from_analysis_code/manuscript_plots_folder",
                                   output_date)

if( !dir.exists( file.path(manuscript_plots_folder) ) ) {
  dir.create( file.path(manuscript_plots_folder), recursive = TRUE )
  
}

supporting_info_plots_folder <- file.path(indicators_project, 
                                     "/Serengeti/Outputs_from_analysis_code/supporting_info_plots_folder",
                                   output_date)

if( !dir.exists( file.path(supporting_info_plots_folder) ) ) {
  dir.create( file.path(supporting_info_plots_folder), recursive = TRUE )
  
}



# Load indicator > scenario level data ----

indicators_all_df <- readRDS(file.path(analysis_inputs, input_date, "all_indicators_output_dataframe.rds"))

indicators_all_list <- flatten(readRDS(file.path(analysis_inputs,input_date, "all_indicators_output_list.rds")))

# Get list of indicators

indicators <- unique(indicators_all_df$indicator)

# Get list of scenarios

scenarios <- unique(indicators_all_df$scenario)

# Save a copy of the inputs used to create outputs

write.csv(indicators_all_df, file.path(analysis_inputs_folder,
                                       paste(output_date, "input_data.csv",
                                             sep = "_")))
# Load red list data

scenario_redlist_data_sampled <- readRDS(file.path(analysis_inputs, input_date, "scenario_redlist_data_annual_7.rds"))
```


```{r prep data, results = FALSE}
# Standardize & difference the data as per Bestelmeyer 2011

indicators_all_df <- indicators_all_df %>% 
                     group_by(indicator, scenario) %>% 
  filter(indicator == "RLI annual"|
         indicator == "LPI"|
         indicator == "total abundance harvested") %>% 
                     mutate(indicator_score_std = standardize(indicator_score),
                            indicator_score_diff = indicator_score_std - 
                             lag(indicator_score_std, default = 0)) %>% 
                     mutate(scenario_indicator = paste(scenario, indicator, 
                                                       sep = "_"),
                            variance = rollapply(indicator_score_diff,
                               10,
                               var,
                               na.rm = TRUE,
                               partial = TRUE,
                               align = "left"),
                            disturbance = as.integer(ifelse(annual_time_step > 99 &
                                                 annual_time_step < 201,
                                                 1, 0)),
                            disturbance_factor = as.factor(ifelse(disturbance == 1,
                                                 "disturbance", 
                                                 "no disturbance")),
                            years_since_disturbance = ifelse(annual_time_step < 100,
                                                            1, (annual_time_step - 100)),
                            disturbance_phase = as.integer(ifelse(annual_time_step < 100,
                                                                  1,
                                                 ifelse(annual_time_step > 99 &
                                                 annual_time_step < 201,
                                                 2, 3)))) %>%
                     ungroup(.)

# Split into a list of scenarios then indicators

scenario_indicators <- split(indicators_all_df, indicators_all_df$scenario)

# Or vice versa

indicator_scenarios <- split(indicators_all_df, indicators_all_df$indicator)

# Single indicator, single scenario

scenario_indicator_single <- split(indicators_all_df, indicators_all_df$scenario_indicator)

# Make a test df of one scenario, one indicator for building loop content

test_df <- indicators_all_df %>% 
           filter(indicator == "total abundance harvested") %>% 
           filter(scenario != "000_Baseline")

# Get the ylimits for indicator plots

# Specify y - limits of indicators

df <- indicators_all_df %>% 
      dplyr::filter(indicator  == "RLI annual") %>% 
      filter(scenario != "000_Baseline")

rlmin <- min(df$indicator_score) 
rlmax <- max(df$indicator_score)
rli_lims <- c(rlmin,rlmax)

df <- indicators_all_df %>% 
      dplyr::filter(indicator  == "LPI") %>% 
      filter(scenario != "000_Baseline")

lpmin <- min(df$indicator_score)
lpmax <- max(df$indicator_score)
lpi_lims <- c(lpmin,lpmax)

df <- indicators_all_df %>% 
      dplyr::filter(indicator  == "total abundance harvested") %>% 
      filter(scenario != "000_Baseline")

hamin <- min(df$indicator_score)
hamax <- max(df$indicator_score)
harv_lims <- c(hamin,hamax)

```

# Baseline test gams
## LPI

First derivatives show the rate of change in the indicator at a specific time
point. Values = 0 signify no trend, values > 0 signify a positive trend, and
values < 0 signify a negative trend.

"The second derivative of the trend curve I(t) at time t is a measure of the curvature of I at that time. If the second derivative is greater than zero, the curve is turning upward. If it is less than zero, the curve is turning downward. The magnitude of the second derivative signifies the tightness of the curvature. Values of approx 0 indicate the curve is linear and changing at a steady state. Years in which the second derivative is markedly different from zero are those in which something is happening to alter the rate of population change." (Fewster et al 2000)

```{r baseline lpi, warnings = FALSE, message = FALSE }

selected_mods <- list()
selected_mods_data <- list()
robust_trends <- list()
deviance_explained <- list()

# Make first three items NULL (first 3 elements are taken up by baseline data 
# which we aren't analysing here)

# robust_trends[[1]] <- NULL
# robust_trends[[2]] <- NULL
# robust_trends[[3]] <- NULL
# 
# deviance_explained[[1]] <- NULL
# deviance_explained[[2]] <- NULL
# deviance_explained[[3]] <- NULL

i <- 1

scenario <- "Baseline"

kvals <- c(10,25,50,100)

ind_method <- "GCV.Cp"

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
         select = TRUE,
         method = ind_method,
         data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

# Extract observed, predicted, derivatives and plot for each test gam

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

residual_plots

lpi_baseline_test_plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

lpi_baseline_test_plots

robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## RLI

```{r baseline rli, message=FALSE, warnings=FALSE}

i <- 2

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
         select = TRUE,
         method = ind_method,
         data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

rli_baseline_test_plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

rli_baseline_test_plots



robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## Harvested group

```{r baseline harvesed, warnings = FALSE, message = FALSE}

i <- 3

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

harvested_baseline_test_plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

harvested_baseline_test_plots

robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

# Land use test gams
## LPI

```{r land use lpi, warnings = FALSE, message = FALSE }

# Make first three items NULL (first 3 elements are taken up by baseline data 
# which we aren't analysing here)

# robust_trends[[1]] <- NULL
# robust_trends[[2]] <- NULL
# robust_trends[[3]] <- NULL
# 
# deviance_explained[[1]] <- NULL
# deviance_explained[[2]] <- NULL
# deviance_explained[[3]] <- NULL

i <- 4 


scenario <- "Land use"

kvals <- c(10,25,50,100)

ind_method <- "GCV.Cp"

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
         select = TRUE,
         method = ind_method,
         data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

# Extract observed, predicted, derivatives and plot for each test gam

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

residual_plots

lpi_landuse_test_plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

lpi_landuse_test_plots

robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## RLI

```{r land use rli, message=FALSE, warnings=FALSE}

i <- 5

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
         select = TRUE,
         method = ind_method,
         data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

rli_landuse_test_plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

rli_landuse_test_plots



robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## Harvested group

```{r harvested rli, warnings = FALSE, message = FALSE}

i <- 6

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

harvested_landuse_test_plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

harvested_landuse_test_plots

robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```


# Harvesting carnivores test gams
## LPI

```{r carnivores lpi, warnings = FALSE, message = FALSE }

i <- 7

scenario <- "Harvesting carnivores"


m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

lpi_carnivore_test_plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

lpi_carnivore_test_plots



robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## RLI

```{r carnivores rli, warnings = FALSE, message = FALSE }

i <- 8 

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

rli_carnivores_test_plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

rli_carnivores_test_plots


robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## Harvested group

```{r carnivores harvested, warnings = FALSE, message = FALSE }

i <- 9

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

harvested_carnivores_test_plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

harvested_carnivores_test_plots


robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

# Harvesting herbivores test gams
## LPI

```{r herbivores lpi, warnings = FALSE }

i <- 10

scenario <- "Harvesting herbivores"

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

lpi_herbivores_test_plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

lpi_herbivores_test_plots


robust_trends[[i]] <- compare_trends(test_gam_data)

trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## RLI

```{r herbivores rli, warnings = FALSE }

i <- 11

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

residual_plots

rli_herbivores_test_plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

rli_herbivores_test_plots


robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)


selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## Harvested group

```{r herbivores harvested, warnings = FALSE }

i <- 12 

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

residual_plots

harvested_carnivores_test_plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

harvested_carnivores_test_plots


robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)


selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

# Baseline selected model
## LPI

```{r model baseline lpi, warnings = FALSE }

bootstrapped_data <- list()
bootstrapped_plots <- list()

i <- 1

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]], 
                                                   lpi_lims)


scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
       bootstrapped_plots[[i]])

bootstrapped_plots[[i]]

```

## RLI

```{r model baseline rli, warnings = FALSE }

i <- 2

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots)%>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   rli_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]


```

## Harvested group

```{r model baseline harvested, warnings = FALSE }

i <- 3

df <- indicators_all_df %>% 
      dplyr::filter(indicator  == "total abundance harvested") %>% 
      filter(scenario == "000_Baseline")

hmin <- min(df$indicator_score)
hmax <- max(df$indicator_score)

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   c(hmin, hmax))

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))


ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]



```

# Land use selected model
## LPI

```{r model land use lpi, warnings = FALSE }

bootstrapped_data <- list()
bootstrapped_plots <- list()

i <- 4

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]], 
                                                   lpi_lims)


scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
       bootstrapped_plots[[i]])

bootstrapped_plots[[i]]

```

## RLI

```{r model land use rli, warnings = FALSE }

i <- 5

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots)%>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   rli_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]


```

## Harvested group

```{r model land use harvested, warnings = FALSE }

i <- 6

# hmax <- get_max(scenario_indicator_single[[i]], "total_abundance_harvested")
# hmin <- get_min(scenario_indicator_single[[i]], "total_abundance_harvested")

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   harv_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))


ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]



```

# Harvesting carnivores selected model
## LPI

```{r model carnivores lpi, warnings = FALSE }

i <- 7

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   lpi_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]


```

## RLI

```{r model carnivores rli, warnings = FALSE }

i <- 8 

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   rli_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]

```

## Harvested group

```{r model carnivores harvested, warnings = FALSE }

i <- 9

df <- indicators_all_df %>% 
      dplyr::filter(indicator  == "total abundance harvested") %>% 
      filter(scenario == "200_Harvesting_carnivores")

hmin <- min(df$indicator_score)
hmax <- max(df$indicator_score)


bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   c(hmin,hmax))

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]


```

# Harvesting herbivores selected models
## LPI

```{r model herbivores lpi, warnings = FALSE }

i <- 10

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   lpi_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))


ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]

```

## RLI

```{r model herbivores rli, warnings = FALSE }

i <- 11

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   rli_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]

```

## Harvested group

```{r model herbivores harvested, warnings = FALSE }

i <- 12 

# hmax <- get_max(scenario_indicator_single[[i]], "total_abundance_harvested")
# hmin <- get_min(scenario_indicator_single[[i]], "total_abundance_harvested")

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   harv_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]


```

# Get periods of change

```{r periods of change, results=FALSE }

all_boot_data <- do.call(rbind, bootstrapped_data) %>% 
                 dplyr::select(annual_time_step, scenario, indicator,
                               robust_positive_trend, robust_negative_trend) %>%
                 filter_at(vars(robust_positive_trend,
                                robust_negative_trend),any_vars(!is.na(.)))
test <- all_boot_data %>% 
        filter(scenario == "100_Land_use") %>% 
        pivot_wider(names_from = indicator,
                    values_from = robust_positive_trend)



```
<!-- # Using MGCV All models -->

<!-- ```{r m all, echo = TRUE} -->

<!-- selected_mods <- list() -->

<!-- for(i in seq_along(scenario_indicator_single)) { -->

<!-- print(scenario_indicator_single[[i]]$scenario_indicator[1]) -->

<!-- selected_mods[[i]] <- gamm(indicator_score ~ s(annual_time_step), -->
<!--           select = TRUE, -->
<!--           data = scenario_indicator_single[[i]], -->
<!--           correlation = corCAR1(form = ~ 1 | annual_time_step)) -->

<!-- } -->

<!-- names(selected_mods) <- names(scenario_indicator_single) -->

<!-- gam.check(selected_mods[[1]]$gam) -->
<!-- names(selected_mods)[1] #LPI -->
<!-- gam.check(selected_mods[[2]]$gam) -->
<!-- names(selected_mods)[2] # RLI -->
<!-- gam.check(selected_mods[[3]]$gam)  -->
<!-- names(selected_mods)[3] # baseline abundance, doesn't matter if weird  -->
<!-- gam.check(selected_mods[[4]]$gam) -->
<!-- names(selected_mods)[4] #LPI -->
<!-- gam.check(selected_mods[[5]]$gam) -->
<!-- names(selected_mods)[5] #RLI -->
<!-- gam.check(selected_mods[[6]]$gam)  -->
<!-- names(selected_mods)[6] # total abundance harvested, may not matter either -->
<!-- gam.check(selected_mods[[7]]$gam) -->
<!-- names(selected_mods)[7] #LPI -->
<!-- gam.check(selected_mods[[8]]$gam)  -->
<!-- names(selected_mods)[8] # RLI annual carnivores, IMPORTANT, needs more work -->
<!-- gam.check(selected_mods[[9]]$gam)  -->
<!-- names(selected_mods)[9] # Abundance carnivores, may not matter -->
<!-- gam.check(selected_mods[[10]]$gam)  -->
<!-- names(selected_mods)[10] # LPI herbivores, IMPORTANT, needs more work -->
<!-- gam.check(selected_mods[[11]]$gam) -->
<!-- names(selected_mods)[11] # RLI -->
<!-- gam.check(selected_mods[[12]]$gam)  -->
<!-- names(selected_mods)[12] # Abundance harvested, may not matter -->

<!-- ``` -->




<!-- # Plot fitted against observed -->

<!-- ```{r model selection 1, echo = TRUE } -->

<!-- fitted_observed_plots <- list() -->

<!-- for (i in seq_along(scenario_indicator_single)) { -->

<!-- fit_obs_df <- data.frame(annual_time_step = scenario_indicator_single[[i]]$annual_time_step,  -->
<!--                    fitted = selected_mods[[i]]$gam$fitted.values,  -->
<!--                    observed = scenario_indicator_single[[i]]$indicator_score) -->

<!-- fitted_observed_plots[[i]] <- ggplot() +  -->
<!--   geom_line(data = fit_obs_df, aes(annual_time_step, fitted)) +  -->
<!--   geom_point(data = fit_obs_df, aes(annual_time_step, observed)) -->

<!-- } -->

<!-- fitted_observed_plots[[1]] -->
<!-- plot_trend(selected_mods2[[1]]) -->
<!-- fitted_observed_plots[[2]] -->
<!-- fitted_observed_plots[[3]] -->
<!-- fitted_observed_plots[[4]] -->
<!-- fitted_observed_plots[[5]] -->
<!-- fitted_observed_plots[[6]] -->
<!-- fitted_observed_plots[[7]] -->
<!-- fitted_observed_plots[[8]] -->
<!-- fitted_observed_plots[[9]] -->
<!-- fitted_observed_plots[[10]] -->
<!-- fitted_observed_plots[[11]] -->
<!-- fitted_observed_plots[[12]] -->

<!-- ``` -->

<!-- # Graveyard  -->
<!-- ```{r graveyard, echo = TRUE } -->

<!-- # Using Jess' method -->

<!-- test_plots <- list() -->

<!-- d <- 2 -->

<!-- for (i in seq_along(scenario_indicator_single)) { -->

<!-- fit_obs_df <- data.frame(annual_time_step = scenario_indicator_single[[i]]$annual_time_step,  -->
<!--                    fitted = selected_mods[[i]]$gam$fitted.values,  -->
<!--                    observed = scenario_indicator_single[[i]]$indicator_score_std) -->

<!-- deriv_func <- splinefun(x = fit_obs_df$annual_time_step, y = fit_obs_df$fitted) -->
<!-- deriv_fitted <- as.data.frame(x = deriv_func(seq(1, 290, 1), deriv = d)) -->

<!-- deriv_df <- cbind(ts = fit_obs_df$annual_time_step,d = deriv_fitted, -->
<!--                   observed = fit_obs_df$observed) -->

<!-- names(deriv_df) <- c("annual_time_step", "deriv", "observed") -->

<!-- test_plots[[i]] <- ggplot() + -->
<!--                    #geom_point(data = deriv_df,aes(annual_time_step, observed)) + -->
<!--                    geom_line(data = deriv_df,aes(annual_time_step, deriv)) -->



<!-- } -->

<!-- i <- i + 1 -->
<!-- test_plots[[i]] -->





<!-- ``` -->

<!-- # 2nd derivatives -->

<!-- Second derivatives show when the indicator values have changed direction. -->
<!-- "The second derivative of the trend curbe I(t) at time t is a measure of the curvature of I at that time. If the second derivative is greater than zero, the curve is turning upward. If it is less than zero, the curve is turning downward. The magnitude of the second derivative signifies the tightness of the curvature. Values of approx 0 indicate the curve is linear and changing at a steady state. Years in which the second derivative is markedly different from zero are those in which something is happening to alter the rate of population change." (Fewster et al 2000) -->

<!-- ```{r derivative 2, echo = TRUE } -->

<!-- deriv_func <- splinefun(x = fit_obs_df$annual_time_step, y = fit_obs_df$fitted) -->
<!-- deriv_fitted <- as.data.frame(x = deriv_func(seq(1, 290, 1), deriv = 2)) -->

<!-- deriv_df <- cbind(ts = fit_obs_df$annual_time_step,d = deriv_fitted) -->

<!-- ggplot(deriv_df, aes(ts, `deriv_func(seq(1, 290, 1), deriv = 2)`))+ -->
<!--   geom_line() -->


<!-- gratia_derivs_2 <- derivatives(selected_mods[[i]], order = 2,  -->
<!--                     n = 290, interval = "confidence") -->

<!-- draw(gratia_derivs_1) -->
<!-- draw(gratia_derivs_2) -->


<!-- ``` -->

<!-- ```{r misc } -->

<!-- m1 <- gamm(response_var ~ s(annual_time_step)+ pressure_var_f,  -->
<!--           data = pr_gam_inputs, -->
<!--            correlation = corARMA(form = ~ 1|seasonal_var, p = 1)) -->


<!-- intervals(m1$lme, which = "var-cov")$corStruct -->

<!-- AIC(m, m0, m1$lme) -->

<!-- selected_mod <- m1 -->

<!-- test <- data.frame(annual_time_step = pr_gam_inputs$annual_time_step,  -->
<!--                    fitted = selected_mod$gam$fitted.values,  -->
<!--                    observed = pr_gam_inputs$response_var) -->

<!-- ggplot() +  -->
<!--   geom_line(data = test, aes(annual_time_step, fitted)) +  -->
<!--   geom_point(data = test, aes(annual_time_step, observed)) -->


<!-- deriv_func <- splinefun(x = test$annual_time_step, y = test$fitted) -->
<!-- deriv_fitted <- as.data.frame(x = deriv_func(seq(1, 290, 1), deriv = 2)) -->

<!-- deriv_df <- cbind(ts = test$annual_time_step,d = deriv_fitted) -->

<!-- ggplot(deriv_df, aes(ts, `deriv_func(seq(1, 290, 1), deriv = 2)`))+ -->
<!--   geom_line() -->

<!-- # https://petolau.github.io/Analyzing-double-seasonal-time-series-with-GAM-in-R/ -->
<!-- # https://fromthebottomoftheheap.net/2014/05/09/modelling-seasonal-data-with-gam/ -->
<!-- # https://asbates.rbind.io/2019/04/19/time-series-with-spline-based-models/ -->
<!-- # https://asbates.rbind.io/2019/05/03/gams-for-time-series/  -->

<!-- fit <- auto.arima(resid(m1$lme, type = "normalized"), -->
<!--                        stationary = TRUE, seasonal = TRUE) -->

<!-- fit$coef -->

<!-- plot(forecast(fit,h=20)) -->
<!-- summary(fit) -->

<!-- plot(pr_gam_inputs$pressure_var_f, resid(m1), xlab = "disturbance", -->
<!-- ylab = "Residuals") -->

<!-- plot(pr_gam_inputs$pressure_var_c, resid(m1), xlab = "disturbance level", -->
<!-- ylab = "Residuals") -->

<!-- plot(pr_gam_inputs$annual_time_step, resid(m1), xlab = "annual time step", -->
<!-- ylab = "Residuals") -->



<!-- selected_mod <- mmmm -->

<!-- gam.check(selected_mod$gam) -->

<!-- E <- residuals(selected_mod, type = "response") -->
<!-- I1 <- !is.na(pr_gam_inputs$response_var) -->
<!-- Efull <- vector(length = length(pr_gam_inputs$response_var)) -->
<!-- Efull <- NA -->
<!-- Efull[I1] <- E -->
<!-- acf(Efull, na.action = na.pass, -->
<!-- main = "Auto-correlation plot for residuals") -->

<!-- acf(Efull, plot = FALSE) -->
<!-- pacf(Efull, plot = FALSE) -->

<!-- ``` -->


<!-- # Model selection -->

<!-- ```{r model selection } -->

<!-- #Using ANOVA (as per https://fromthebottomoftheheap.net/2014/05/09/modelling-seasonal-data-with-gam/) -->

<!-- # anova(m1, m2$lme, m3, m4$lme, m5, m6$lme) -->

<!-- ## Using AIC -->
<!-- mod_sel_table <- as.data.frame(AIC(m1, m2$lme, m3$lme, m4)) -->
<!-- mod_sel_table -->

<!-- # Define selected model -->

<!-- selected_mod <- m4 -->

<!-- summary(selected_mod) -->

<!-- plot(selected_mod, residuals = TRUE, pch = 19, cex = 0.75) -->

<!-- gam.check(selected_mod) -->
<!-- ``` -->

<!-- ```{r plot selected mods, eval = FALSE} -->

<!-- summary(pr_gam_inputs) -->

<!-- pdat <- data.frame(annual_time_step = c(1:300), -->
<!--                    pressure_var_f = as.factor("disturbance")) -->

<!-- head(pdat) -->

<!-- p <- predict(selected_mod, newdata = pdat, type = "response") -->
<!-- p -->

<!-- pdat <- cbind(pdat,p) -->
<!-- lines(p ~ annual_time_step, data = pdat, col = "pink") -->
<!-- # lines(p1 ~ annual_time_step, data = pdat, col = "red") -->
<!-- # lines(p2 ~ annual_time_step, data = pdat, col = "blue") -->
<!-- legend("topleft", -->
<!--        legend = c("Uncorrelated Errors","AR(1) Errors"), -->
<!--        bty = "n", col = c("red","blue"), lty = 1) -->

<!-- # *** Plot the derivatives and periods of change ---- -->

<!-- m2.d <- Deriv(selected_mod$terms, n = 300) -->
<!-- plot(m2.d, sizer = TRUE, alpha = 0.01) -->

<!-- # Add periods of change to time series -->

<!-- plot(LPI ~ annual_time_step, data = pr_gam_inputs,  -->
<!--      type = "p") -->
<!-- lines(p ~ annual_time_step, data = pdat) -->
<!-- CI <- confint(m2.d, alpha = 0.01) -->
<!-- S <- signifD(p, m2.d$annual_time_step$deriv,  -->
<!--              CI$annual_time_step$upper,  -->
<!--              CI$annual_time_step$lower, -->
<!--              eval = 0) -->
<!-- lines(S$incr ~ annual_time_step, data = pdat, lwd = 3, col = "blue") -->
<!-- lines(S$decr ~ annual_time_step, data = pdat, lwd = 3, col = "red") -->
<!-- # Recreate with ggplot -->

<!-- deriv_plot <- ggplot() + -->
<!--   geom_point(data = gam_inputs, -->
<!--              aes(x = annual_time_step, y = indicator_score), -->
<!--              alpha = 0.3) + -->
<!--   geom_line(aes(x = gam_inputs$annual_time_step, -->
<!--                 y = p2)) + -->
<!--   geom_line(aes(x = gam_inputs$annual_time_step,  -->
<!--                 y = S$incr), col = "blue", size = 2) + -->
<!--   geom_line(aes(x = gam_inputs$annual_time_step,  -->
<!--                 y = S$decr), col = "red", size = 2) + -->
<!--   labs(title = paste(gam_inputs$scenario[1], -->
<!--                      gam_inputs$indicator[1], sep = " ")) -->

<!-- deriv_plot -->

<!-- ggsave(file.path(analysis_plots_folder, -->
<!--                  paste(gam_inputs$indicator[1],"_", gam_inputs$scenario[1],  -->
<!--                        "derivative_plot.png", sep = "")), -->
<!--        deriv_plot,  device = "png") -->

<!-- ``` -->
