---
title: "Analyse terrestrial indicators"
author: "Simone Stevenson"
date: "20/09/2021"
output:
  html_document: default
  pdf_document: default
---

#' TODO: Fix axis labels + titles on bootstrapped plots
#' TODO: Add vector art for scenarios (lion, wildebeest, plant)
#' TODO: Add direct/indirect label to spp response plots
#' TODO: Add matched species names on spp gam plots
#' TODO: Add functional group colours to species GAMS

```{r setup, include=FALSE}

# https://bookdown.org/yihui/rmarkdown/r-code.html

knitr::opts_chunk$set(echo = FALSE, message = FALSE,
  warnings = FALSE, error = FALSE)

```

Directory path to git repo

Simone's deakin laptop

cd "C:\\Users\\ssteven\\OneDrive - Deakin University\\Deakin\\Chapter_3_indicator_testing\\madingley_terrestrial_indicators"

Data structure

Input data is structured hierarchically as follows:
- Location (1)
-- Scenarios (3)
--- Replicates (25)

Output indicator data is calculated at the same level, so each indicator should
have values for:
- Location (1)
-- Scenarios (3)
--- Replicates (25)

# Possible useful ref?
https://www.frontiersin.org/articles/10.3389/fevo.2018.00149/full

```{r libraries}

## Data wrangling
library(tidyverse)
library(zoo)

## Analysis
library(mgcv)
library(nlme)
library(MASS)
library(datawizard)

## Viz
library(ggpubr)
library(cowplot)
library(gridExtra)
library(gratia)
library(kableExtra)
library(viridis)
library(viridisLite)

```

```{r template chunk}

```

```{r ggplot theme}

plot_theme <- theme(panel.background = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              legend.position = "bottom",
              legend.key.size = unit(0.25, 'cm'),
              plot.title = element_text(size = 10, face = "bold"),
              legend.title=element_text(size=6), 
              legend.text=element_text(size=5),
              axis.line = element_line(colour = "black"),
              axis.title.x = element_text(size=8),
              axis.title.y = element_text(size=8),
              axis.text.x = element_text(size=8),
              axis.text.y = element_text(size=8),
              axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank())

```

```{r colour palettes, eval=FALSE}

library(scales)
show_col(viridis_pal()(20))
show_col(viridis_pal(option = "plasma")(20))
show_col(viridis_pal(option = "magma")(20))
show_col(viridis_pal(option = "inferno")(20))
show_col(viridis_pal(option = "mako")(20))
show_col(viridis_pal(option = "cividis")(20))
show_col(viridis_pal(option = "rocket")(20))
show_col(viridis_pal(option = "turbo")(20))


```

```{r get species gam data }

#' @param model - gam object
#' @param data - the input pressure response time step data for the model
#' @return a dataframe containing predicted, observed, first and second derivatives

get_species_gam_data <- function(data, model, scenario) {
  
# Get observed indicator values

observed <- data$std_abundance

# Make a dataframe of predictor vals (ie timesteps)

pdat1 <- data.frame(annual_time_step = seq(1,nrow(data),1))

# Get predicted values for each time step
# p2 <- predict.gam(model, pdat1, 
#                   se.fit = TRUE, response = TRUE)

p2 <- predict(model, pdat1, 
                  se.fit = TRUE, response = TRUE)

p3 <- predict.gam(model, pdat1, 
                  se.fit = TRUE)

# # Get the deviance residuals
deviance_residuals <- residuals(model, type = "working")

# fv <- predict(model,type="terms") ## get term estimates
#      
# ## compute partial residuals for first smooth...          
# deviance_residuals <- residuals(model,type="working", response = TRUE) + fv[,1]

#Bind them together with observed vals

pdat <- cbind(pdat1, predicted = p2$fit, predicted_nr = p3$fit, se2 = p2$se.fit, observed, 
              deviance_residuals)

#Get upper and lower bounds for predicted values (replace with bootstraps for selected model)

df.res <- df.residual(model)
crit.t <- qt(0.025, df.res, lower.tail = FALSE)
pdat <- transform(pdat,
                  upper_pred = predicted_nr + (crit.t * se2),
                  lower_pred = predicted_nr - (crit.t * se2))

# Calculate the derivatives and add that too
derivs_1 <- as.data.frame(derivatives(model, order = 1, 
                   n = nrow(data), interval = "confidence") ) %>% 
                   dplyr::select(derivative, lower, upper) %>% 
                   rename(first_derivative = derivative,
                          first_lower = lower,
                          first_upper = upper)

# Calculate the derivatives and add that too
derivs_2 <- as.data.frame(derivatives(model, order = 2, 
                   n = nrow(data), interval = "confidence") ) %>% 
                   dplyr::select(derivative, lower, upper) %>% 
                   rename(second_derivative = derivative,
                          second_lower = lower,
                          second_upper = upper)
# Bind it all together
temp <- cbind(pdat, derivs_1, derivs_2)


# Make the final plot data
test_gam_data <- temp %>% 
                 mutate(first_significant = ifelse(first_lower <= 0 & 
                                                  first_upper >= 0, FALSE, TRUE),
                        positive_trend = ifelse(first_significant == TRUE & 
                                                 first_derivative > 0,
                                            predicted_nr,NA),
                        negative_trend = ifelse(first_significant == TRUE & 
                                                 first_derivative < 0,
                                            predicted_nr, NA),
                        positive_der_1 = ifelse(first_significant == TRUE & 
                                                  first_derivative > 0,
                                            first_derivative,NA),
                        negative_der_1 = ifelse(first_significant == TRUE & 
                                                  first_derivative < 0,
                                            first_derivative, NA),
                        second_significant = ifelse(second_lower <= 0 & 
                                                  first_upper >= 0, FALSE, TRUE),
                        positive_tp = ifelse(second_significant == TRUE & 
                                             second_derivative > 0,
                                            predicted_nr,NA),
                        negative_tp = ifelse(second_significant == TRUE & 
                                            second_derivative < 0,
                                            predicted_nr, NA),
                        positive_der_2 = ifelse(second_significant == TRUE & 
                                                second_derivative > 0,
                                            second_derivative,NA),
                        negative_der_2 = ifelse(second_significant == TRUE & 
                                                second_derivative < 0,
                                            first_derivative, NA),
                        indicator = data$group_id[1],
                        scenario = scenario,
                        kvalue = model$rank,
                        r2 = summary.gam(model)$r.sq) 

}
```

```{r get test gam data }

#' @param model - gam object
#' @param data - the input pressure response time step data for the model
#' @return a dataframe containing predicted, observed, first and second derivatives

get_test_gam_data <- function(data, model, scenario) {
  
# Get observed indicator values

observed <- data$indicator_score

# Make a dataframe of predictor vals (ie timesteps)

pdat1 <- data.frame(annual_time_step = seq(1,nrow(data),1))

# Get predicted values for each time step
# p2 <- predict.gam(model, pdat1, 
#                   se.fit = TRUE, response = TRUE)

p2 <- predict(model, pdat1, 
                  se.fit = TRUE, response = TRUE)

p3 <- predict.gam(model, pdat1, 
                  se.fit = TRUE)

# # Get the deviance residuals
deviance_residuals <- residuals(model, type = "working")

# fv <- predict(model,type="terms") ## get term estimates
#      
# ## compute partial residuals for first smooth...          
# deviance_residuals <- residuals(model,type="working", response = TRUE) + fv[,1]

#Bind them together with observed vals

pdat <- cbind(pdat1, predicted = p2$fit, predicted_nr = p3$fit, se2 = p2$se.fit, observed, 
              deviance_residuals)

#Get upper and lower bounds for predicted values (replace with bootstraps for selected model)

df.res <- df.residual(model)
crit.t <- qt(0.025, df.res, lower.tail = FALSE)
pdat <- transform(pdat,
                  upper_pred = predicted_nr + (crit.t * se2),
                  lower_pred = predicted_nr - (crit.t * se2))

# Calculate the derivatives and add that too
derivs_1 <- as.data.frame(derivatives(model, order = 1, 
                   n = nrow(data), interval = "confidence") ) %>% 
                   dplyr::select(derivative, lower, upper) %>% 
                   rename(first_derivative = derivative,
                          first_lower = lower,
                          first_upper = upper)

# Calculate the derivatives and add that too
derivs_2 <- as.data.frame(derivatives(model, order = 2, 
                   n = nrow(data), interval = "confidence") ) %>% 
                   dplyr::select(derivative, lower, upper) %>% 
                   rename(second_derivative = derivative,
                          second_lower = lower,
                          second_upper = upper)

# derivs_2_fun <- splinefun(x = data$annual_time_step, y = pdat$predicted)
# 
# derivs_2_2 <- derivs_2_fun(data$annual_time_step, deriv = 2)

# Bind it all together
temp <- cbind(pdat, derivs_1, derivs_2)

# Get the indicator we are modeling
indicator <- data$indicator[1]
#scenario <- data$scenario[1]

# Make the final plot data
test_gam_data <- temp %>% 
                 mutate(first_significant = ifelse(first_lower <= 0 & 
                                                  first_upper >= 0, FALSE, TRUE),
                        positive_trend = ifelse(first_significant == TRUE & 
                                                 first_derivative > 0,
                                            predicted_nr,NA),
                        negative_trend = ifelse(first_significant == TRUE & 
                                                 first_derivative < 0,
                                            predicted_nr, NA),
                        positive_der_1 = ifelse(first_significant == TRUE & 
                                                  first_derivative > 0,
                                            first_derivative,NA),
                        negative_der_1 = ifelse(first_significant == TRUE & 
                                                  first_derivative < 0,
                                            first_derivative, NA),
                        second_significant = ifelse(second_lower <= 0 & 
                                                  first_upper >= 0, FALSE, TRUE),
                        positive_tp = ifelse(second_significant == TRUE & 
                                             second_derivative > 0,
                                            predicted_nr,NA),
                        negative_tp = ifelse(second_significant == TRUE & 
                                            second_derivative < 0,
                                            predicted_nr, NA),
                        positive_der_2 = ifelse(second_significant == TRUE & 
                                                second_derivative > 0,
                                            second_derivative,NA),
                        negative_der_2 = ifelse(second_significant == TRUE & 
                                                second_derivative < 0,
                                            first_derivative, NA),
                        indicator = indicator,
                        scenario = scenario,
                        kvalue = model$rank,
                        r2 = summary.gam(model)$r.sq) 

}
```

```{r plot test gam}

plot_test_gam <- function(gam_data) {

if(all(is.na(gam_data$positive_trend)) & 
           !all(is.na(gam_data$negative_trend))) {

gam_plot <- ggplot(data = gam_data) +
            geom_rect(aes(xmin = 95, xmax = 105,
                          ymin = -Inf, ymax = Inf), fill = "grey82", 
                            alpha = 0.4) +
            geom_rect(aes(xmin = 195, xmax = 205,
                          ymin = -Inf, ymax = Inf), fill = "grey82", 
                            alpha = 0.4)+
            geom_vline(xintercept=c(100,200), linetype = "dotted") +
            #geom_point(aes(x = annual_time_step, y = deviance_residuals), alpha = 0.2) +
            geom_point(aes(x = annual_time_step, y = observed), alpha = 0.2) +
  # geom_point(aes(x = annual_time_step, y = positive_tp), shape = 8, size = 3, col = "blue") +
  # geom_point(aes(x = annual_time_step, y = negative_tp), shape = 8, size = 3, col = "red") +
            geom_rug(sides="b", aes(x = annual_time_step)) +
            # Plot the smoothed indicator line
            geom_line(aes(x = annual_time_step, y = predicted_nr)) +
            # Add the confidence intervals
            geom_ribbon(aes(x = annual_time_step, ymin = lower_pred, ymax = upper_pred),
                        alpha = 0.3) +
            # Add the significant trends based on first derivative
            #geom_line(aes(x = annual_time_step, y = positive_trend), col = "blue") +
            geom_line(aes(x = annual_time_step, y = negative_trend), col = neg_col) +
            labs(y = gam_data$indicator[1],
                 title = paste(gam_data$scenario[1], 
                               gam_data$indicator[1], "k =", 
                               gam_data$kvalue[1], sep = " ")) +
            plot_theme

} else if(!all(is.na(gam_data$positive_trend))& 
           all(is.na(gam_data$negative_trend))) {

gam_plot <- ggplot(data = gam_data) +
  geom_rect(aes(xmin = 95, xmax = 105,
                          ymin = -Inf, ymax = Inf), fill = "grey82", 
                            alpha = 0.4) +
            geom_rect(aes(xmin = 195, xmax = 205,
                          ymin = -Inf, ymax = Inf), fill = "grey82", 
                            alpha = 0.4)+
            geom_vline(xintercept=c(100,200), linetype = "dotted") +
            #geom_point(aes(x = annual_time_step, y = deviance_residuals), alpha = 0.2) +
            geom_point(aes(x = annual_time_step, y = observed), alpha = 0.2) +
            geom_rug(sides="b", aes(x = annual_time_step)) +
            # Plot the smoothed indicator line
            geom_line(aes(x = annual_time_step, y = predicted_nr)) +
            # Add the confidence intervals
            geom_ribbon(aes(x = annual_time_step, ymin = lower_pred, ymax = upper_pred),
                        alpha = 0.3) +
            # Add the significant trends based on first derivative
            geom_line(aes(x = annual_time_step, y = positive_trend), col = pos_col) +
            #geom_line(aes(x = annual_time_step, y = negative_trend), col = "red") +
            labs(y = gam_data$indicator[1],
                 title = paste(gam_data$scenario[1], 
                               gam_data$indicator[1], "k =", 
                               gam_data$kvalue[1], sep = " ")) +
            plot_theme


} else if (all(is.na(gam_data$positive_trend)) & 
           all(is.na(gam_data$negative_trend))) {

gam_plot <- ggplot(data = gam_data) +
  geom_rect(aes(xmin = 95, xmax = 105,
                          ymin = -Inf, ymax = Inf), fill = "grey82", 
                            alpha = 0.4) +
            geom_rect(aes(xmin = 195, xmax = 205,
                          ymin = -Inf, ymax = Inf), fill = "grey82", 
                            alpha = 0.4)+
            geom_vline(xintercept=c(100,200), linetype = "dotted") +
            #geom_point(aes(x = annual_time_step, y = deviance_residuals), alpha = 0.2) +
            geom_point(aes(x = annual_time_step, y = observed), alpha = 0.2) +
            geom_rug(sides="b", aes(x = annual_time_step)) +
            # Plot the smoothed indicator line
            geom_line(aes(x = annual_time_step, y = predicted_nr)) +
            # Add the confidence intervals
            geom_ribbon(aes(x = annual_time_step, ymin = lower_pred, ymax = upper_pred),
                        alpha = 0.3) +
            # Add the significant trends based on first derivative
            #geom_line(aes(x = annual_time_step, y = positive_trend), col = "blue") +
            #geom_line(aes(x = annual_time_step, y = negative_trend), col = "red") +
            labs(y = gam_data$indicator[1],
                 title = paste(gam_data$scenario[1], 
                               gam_data$indicator[1], "k =", 
                               gam_data$kvalue[1], sep = " ")) +
            plot_theme


} else {
  
gam_plot <- ggplot(data = gam_data) +
  geom_rect(aes(xmin = 95, xmax = 105,
                          ymin = -Inf, ymax = Inf), fill = "grey82", 
                            alpha = 0.4) +
            geom_rect(aes(xmin = 195, xmax = 205,
                          ymin = -Inf, ymax = Inf), fill = "grey82", 
                            alpha = 0.4)+
            geom_vline(xintercept=c(100,200), linetype = "dotted") +
            #geom_point(aes(x = annual_time_step, y = deviance_residuals), alpha = 0.2) +
            geom_point(aes(x = annual_time_step, y = observed), alpha = 0.2) +
            geom_rug(sides="b", aes(x = annual_time_step)) +
            # Plot the smoothed indicator line
            geom_line(aes(x = annual_time_step, y = predicted_nr)) +
            # Add the confidence intervals
            geom_ribbon(aes(x = annual_time_step, ymin = lower_pred, ymax = upper_pred),
                        alpha = 0.3) +
            # Add the significant trends based on first derivative
            geom_line(aes(x = annual_time_step, y = positive_trend), col = pos_col) +
            geom_line(aes(x = annual_time_step, y = negative_trend), col = neg_col) +
            labs(y = gam_data$indicator[1],
                 title = paste(gam_data$scenario[1], 
                               gam_data$indicator[1], "k =", 
                               gam_data$kvalue[1], sep = " ")) +
            plot_theme

return(gam_plot)

  }
}


```

```{r compare trends}
# Loop through each model and identify which time steps show a significant 
# positive or negative trend
# list_of_model_data <- test_gam_data

compare_trends <- function(list_of_model_data) {

# Get the number of models we need consensus across
l <- length(list_of_model_data)

trend_data <- list()

for (j in seq_along(list_of_model_data)) {
  
trend_data[[j]] <- list_of_model_data[[j]] %>% 
                   dplyr::select(annual_time_step, positive_trend,
                              negative_trend, indicator, scenario,
                              first_derivative, kvalue) %>% 
                   mutate(k_value = kvals[[j]],
                          direction = ifelse(!is.na(positive_trend),
                                             "positive",
                                            ifelse(!is.na(negative_trend),
                                             "negative", NA)))
}

# Convert  into a single data frame
trend_df_all <- do.call(rbind, trend_data)


# Check which time steps show a significant trend regardless of model
# k value (ie are the robust significant changes at the expected time points?)

# direction_df <- trend_df_all %>% 
#                 dplyr::select(annual_time_step, k_value, direction, positive_trend,
#                               negative_trend, indicator, scenario, first_derivative) %>% 
#                 group_by(annual_time_step) %>% 
#                 mutate(positive_across_all_models = sum(direction == "positive", na.rm = TRUE),
#                        negative_across_all_models = sum(direction == "negative", na.rm = TRUE)) %>% 
#                 filter(positive_across_all_models == l|
#                       negative_across_all_models == l) #%>% 
#                 # mutate(positive_across_all_models = ifelse(positive_across_all_models == l, 1, 0),
                #        negative_across_all_models = ifelse(negative_across_all_models == l, 1, 0))



# Check which time steps show a significant trend regardless of model
# k value (ie are the robust significant changes at the expected time points?)
direction_df <- trend_df_all %>% 
                dplyr::select(annual_time_step, k_value, direction, positive_trend,
                              negative_trend, indicator, scenario, first_derivative) %>% 
                group_by(annual_time_step) %>% 
                summarise(positive_across_all_models = sum(direction == "positive"),
                          negative_across_all_models = sum(direction == "negative")) %>% 
                filter(positive_across_all_models == l|
                       negative_across_all_models == l) #%>% 
                # mutate(positive_across_all_models = ifelse(positive_across_all_models == l, 1, 0),
                #        negative_across_all_models = ifelse(negative_across_all_models == l, 1, 0))

return(direction_df)

}

```

```{r bootstrap selected model}

bootstrap_selected_model <- function(data, model, model_data, numboots) {

bootstrap_mods <- list()
bootstrap_mod_data <- list()
replicate <- list()

for (k in 1:numboots) {
        
# Take k number of random samples from the input data
replicate[[k]] <- slice_sample(data, 
                          prop = 1, replace = TRUE) 
        

# Fit the model with the data
bootstrap_mods[[k]] <- gam(indicator_score ~ s(annual_time_step, k = model$rank),
          select = TRUE,
          method = "GCV.Cp",
          data = replicate[[k]])

# Get the predicted values etc

obs <- replicate[[k]]$indicator_score

# Make a dataframe of predictor vals (ie timesteps)

pdat1 <- data.frame(annual_time_step = seq(1,nrow(data),1))

# Get predicted values for each time step
p2 <- predict.gam(bootstrap_mods[[k]], pdat1,
                  se.fit = TRUE, response = TRUE)

# Get the deviance residuals
deviance_residuals <- residuals(bootstrap_mods[[k]], type = "deviance", 
                                response = TRUE)

#Bind them together with observed vals

pdat <- cbind(pdat1, predicted = p2$fit, se2 = p2$se.fit, obs, 
              deviance_residuals) %>%  
                  mutate(replicate = k)

# Calculate the derivatives and add that too
gratia_derivs_1 <- as.data.frame(derivatives(bootstrap_mods[[k]], order = 1, 
                   n = nrow(data), interval = "confidence") ) %>% 
                   dplyr::select(derivative, lower, upper) %>% 
                   rename(first_derivative = derivative,
                          first_lower = lower,
                          first_upper = upper)

# Calculate the derivatives and add that too
gratia_derivs_2 <- as.data.frame(derivatives(bootstrap_mods[[k]], order = 2, 
                   n = nrow(data), interval = "confidence") ) %>% 
                   dplyr::select(derivative, lower, upper) %>% 
                   rename(second_derivative = derivative,
                          second_lower = lower,
                          second_upper = upper)

# Bind it all together
bootstrap_mod_data[[k]] <- cbind(pdat, gratia_derivs_1, gratia_derivs_2)

print(k)

}

ci_df <- do.call(rbind, bootstrap_mod_data) %>% 
             #Group by timestep
             group_by(annual_time_step) %>% 
             arrange(predicted) %>% 
             summarise(predicted_lower = quantile(predicted, 
                                      probs = 0.025),
                       predicted_upper = quantile(predicted, 
                                      probs = 0.975)) 

first_deriv_df <- do.call(rbind, bootstrap_mod_data) %>% 
             #Group by timestep
             group_by(annual_time_step) %>% 
             # Arrange first deriv values from each bootstrap in order
             arrange(first_derivative) %>% 
             # Set bottom 25th and top 75th quantiles as the lower and upper CIs of the first derivative
             summarise(first_lower = quantile(first_derivative, 
                                      probs = 0.025),
                       first_upper = quantile(first_derivative, 
                                      probs = 0.975)) %>% 
             # Add in the original first derivative from the full dataset
             merge(model_data[c("annual_time_step", "predicted",
                                "first_derivative")],
                   by = "annual_time_step") %>% 
             # Check if the derivative CIs include zero (if so trend is not significant)
             mutate(significant = ifelse(first_lower < 0 & first_upper > 0, 
                                         FALSE, TRUE),
             # If trend is significant and first deriv > 0, get the the predicted indicator 
             # trend values that are positive
                    positive_trend = ifelse(significant == TRUE & first_derivative > 0,
                                            predicted,NA),
             # If trend is significant and first deriv > 0, get the the predicted indicator 
             # trend values that are negative
                    negative_trend = ifelse(significant == TRUE & first_derivative < 0,
                                            predicted, NA),
             # If trend is significant and first deriv > 0, get the the predicted first 
             # derivative values that are positive
                    positive_der1 = ifelse(significant == TRUE & first_derivative > 0,
                                            first_derivative,NA),
             # If trend is significant and first deriv > 0, get the the predicted first 
             # derivative values that are negative
                    negative_der1 = ifelse(significant == TRUE & first_derivative < 0,
                                            first_derivative, NA)) %>% 
            dplyr::select(-predicted)

second_deriv_df <- do.call(rbind, bootstrap_mod_data) %>% 
             #Group by timestep
             group_by(annual_time_step) %>% 
             # Order second derivative values 
             arrange(second_derivative) %>% 
             summarise(second_lower = quantile(second_derivative, 
                                      probs = 0.025),
                       second_upper = quantile(second_derivative, 
                                      probs = 0.975))%>% 
             # Add in the original first derivative from the full dataset
             merge(model_data[c("annual_time_step", "predicted",
                                "second_derivative")],
                   by = "annual_time_step") %>% 
             mutate(significant_2 = ifelse(second_lower < 0 & second_upper > 0, 
                                         FALSE, TRUE),
                    positive_tp = ifelse(significant_2 == TRUE & second_derivative > 0,
                                            predicted,NA),
                    negative_tp = ifelse(significant_2 == TRUE & second_derivative < 0,
                                            predicted, NA),
                    positive_der2 = ifelse(significant_2 == TRUE & second_derivative > 0,
                                            second_derivative,NA),
                    negative_der2 = ifelse(significant_2 == TRUE & second_derivative < 0,
                                            second_derivative, NA)) %>% 
            dplyr::select(-predicted)


new <- model_data %>% 
     dplyr::select(annual_time_step, observed, predicted, deviance_residuals, 
                   indicator, scenario, kvalue) %>% 
     merge(ci_df, by ="annual_time_step") %>% 
     merge(first_deriv_df, by ="annual_time_step") %>% 
     merge(second_deriv_df, by ="annual_time_step") #%>% 
    # merge(model_data[c("annual_time_step", "kvalue")], by = "annual_time_step")

}


```

```{r plot bootstrapped model, eval = FALSE}

plot_bootstrapped_model <- function(data, ylimits){
  
indicator <- data$indicator[1]

boot_base_plot <- ggplot(data = data) +
                  geom_rect(aes(xmin = 95, xmax = 105,
                                ymin = -Inf, ymax = Inf), fill = "grey82", 
                            alpha = 0.4) +
                  geom_rect(aes(xmin = 195, xmax = 205,
                                ymin = -Inf, ymax = Inf), fill = "grey82", 
                            alpha = 0.4)+
                  geom_point(aes(x = annual_time_step, y = observed), 
                                       alpha = 0.2, size = 1) +
                  # geom_line(aes(x = annual_time_step, y = observed), 
                  #           alpha = 0.2) +
                  geom_line(aes(x = annual_time_step, y = predicted)) +
                  geom_ribbon(aes(x = annual_time_step, ymin = predicted_lower, 
                                            ymax = predicted_upper),
                                        alpha = 0.4) +
                  labs(x = "Time (years)",
                        y = indicator) + 
                  geom_vline(xintercept=c(100,200), linetype = "dotted") +
                  coord_cartesian(ylim = ylimits)
                # + annotate("text", x = 295, y = 1.1, 
                #            label = paste("Dev. explained =",
                #                          dev_explained,sep = " "), 
                #            size = 4)
if(all(is.na(data$positive_trend))) {
boot_plot <-  boot_base_plot +
              # geom_line(aes(x = annual_time_step, y = negative_trend), 
              #           col = neg_col, size = 1) +
              plot_theme 
  
} else if(all(is.na(data$robust_positive_trend))) {
boot_plot <- boot_base_plot +
  #  geom_line(aes(x = annual_time_step, y = positive_trend), col = pos_col, 
  #            size = 1) +
  # geom_line(aes(x = annual_time_step, y = negative_trend), col = neg_col, 
  #            size = 1) +
  geom_line(aes(x = annual_time_step, y = robust_negative_trend), 
            col = neg_col, size = 4) +
  plot_theme  
 
  
} else if (all(is.na(data$robust_negative_trend))) {
boot_plot <- boot_base_plot +
  #  geom_line(aes(x = annual_time_step, y = positive_trend), col = pos_col, 
  #            size = 1) +
  # geom_line(aes(x = annual_time_step, y = negative_trend), col = neg_col, 
  #            size = 1) +
  geom_line(aes(x = annual_time_step, y = robust_positive_trend), 
            col = pos_col, size = 4) +
  plot_theme 
 
} else {
  boot_plot <- boot_base_plot +
  # geom_line(aes(x = annual_time_step, y = positive_trend), col = pos_col, 
  #            size = 1) +
  # geom_line(aes(x = annual_time_step, y = negative_trend), col = neg_col, 
  #            size = 1) +
  geom_line(aes(x = annual_time_step, y = robust_positive_trend), 
            col = pos_col, size = 4) +
  geom_line(aes(x = annual_time_step, y = robust_negative_trend), 
            col = neg_col, size = 4) +
  plot_theme 
  
  return(boot_plot)
  
  }
}
```

# Inputs

```{r data, results=FALSE }

# Set up paths ----

location <- 'Serengeti'

indicators_project <- "N:/Quantitative-Ecology/Indicators-Project"

# Analysis inputs are the outputs of 2_calculate_madingley_indicators.R

analysis_inputs <- "N:/Quantitative-Ecology/Indicators-Project/Serengeti/Outputs_from_indicator_code/Indicator_outputs/general/"

# Inputs ----

# Specify the date of inputs you want to use

input_date <- "2021-09-30"

# Specify how many bootstraps you want to use

boots <- 2

# Specify how many influential species you want to plot

n <- 6

# Size of test gam plots

t_width <- 20
t_height <- 10

# Make colours for pos and neg trends

# Purple and green?
# neg_col <- "#440154FF"
# pos_col <- "#2A788EFF"

# Pink and purple
# neg_col <- "#C5407EFF"
# pos_col <- "#2D0594FF"

# Red and blue
neg_col <- "#A51301FF"
pos_col <- "#3AA2FCFF"

# Get date to label outputs

output_date <- Sys.Date() 

scenarios_title <- list("Baseline", "Land use", 
                  "Carnivore harvesting", 
                  "Herbivore harvesting")

scenarios_title_2 <- c("Baseline","Baseline","Harvesting primary producers","Harvesting primary producers",
                       "Harvesting carnivores","Harvesting carnivores",
                       "Harvesting herbivores", "Harvesting herbivores")

disturbance_string <- c("pre-disturbance", "disturbance", "post-disturbance")

disturbance_factor <- factor(disturbance_string, ordered = TRUE,
                              levels = c("pre-disturbance", 
                                         "disturbance", 
                                         "post-disturbance"))

# Set up output folders

## Include a folder to store the version of inputs used to get the outputs

analysis_inputs_folder <- file.path(indicators_project,  
                           "/Serengeti/Outputs_from_analysis_code/Analysis_inputs",
                           output_date)

if( !dir.exists( file.path(analysis_inputs_folder) ) ) {
  dir.create( file.path(analysis_inputs_folder), recursive = TRUE )
  
}

analysis_outputs_folder <- file.path(indicators_project, 
                                     "/Serengeti/Outputs_from_analysis_code/Analysis_outputs",
                                     output_date)

if( !dir.exists( file.path(analysis_outputs_folder) ) ) {
  dir.create( file.path(analysis_outputs_folder), recursive = TRUE )
  
}

manuscript_plots_folder <- file.path(indicators_project, 
                                     "/Serengeti/Outputs_from_analysis_code/manuscript_plots_folder",
                                   output_date)

if( !dir.exists( file.path(manuscript_plots_folder) ) ) {
  dir.create( file.path(manuscript_plots_folder), recursive = TRUE )
  
}

supporting_info_plots_folder <- file.path(indicators_project, 
                                     "/Serengeti/Outputs_from_analysis_code/supporting_info_plots_folder",
                                   output_date)

if( !dir.exists( file.path(supporting_info_plots_folder) ) ) {
  dir.create( file.path(supporting_info_plots_folder), recursive = TRUE )
  
}



# Load indicator > scenario level data ----

indicators_all_df <- readRDS(file.path(analysis_inputs, input_date, "all_indicators_output_dataframe.rds"))

indicators_all_list <- flatten(readRDS(file.path(analysis_inputs,input_date, "all_indicators_output_list.rds")))

# Get list of indicators

indicators <- unique(indicators_all_df$indicator)

# Get list of scenarios

scenarios <- unique(indicators_all_df$scenario)

# Save a copy of the inputs used to create outputs

write.csv(indicators_all_df, file.path(analysis_inputs_folder,
                                       paste(output_date, "input_data.csv",
                                             sep = "_")))
# Load red list data

scenario_redlist_data_sampled <- readRDS(file.path(analysis_inputs, input_date, "scenario_redlist_data_annual_7.rds"))

# Load autotroph data

auto_annual <- readRDS(file.path(analysis_inputs, input_date, 
                                 "scenario_auto_annual_6.rds"))
# Get groups

groups <- read.csv(file.path(analysis_inputs, input_date, "groups.csv"))

# Add 'harvested' tag

groups <- groups %>% 
          mutate(harvested = ifelse(group_id == "10.67", "(directly harvested)",
                             ifelse(group_id == "10.68", "(directly harvested)",
                             ifelse(group_id == "10.69", "(directly harvested)",
                             ifelse(group_id == "11.67", "(directly harvested)",
                             ifelse(group_id == "11.68", "(directly harvested)",
                             ifelse(group_id == "11.69", "(directly harvested)", 
                                    " ")))))))

species <- read.csv("N:/Quantitative-Ecology/Indicators-Project/Serengeti/Outputs_from_indicator_code/Indicator_outputs/general/2021-09-22/model_species.csv")

# Get 1 spp per group
example_species <- species %>% 
                   filter(!is.na(common_name)) %>% 
                   group_by(group_id) %>% 
                   slice(1)

```


```{r prep data, results = FALSE}
# Standardize & difference the data as per Bestelmeyer 2011

indicators_all_df <- indicators_all_df %>% 
                     group_by(indicator, scenario) %>% 
  filter(indicator == "RLI annual"|
         indicator == "LPI") %>% 
                     mutate(indicator_score_std = standardize(indicator_score),
                            indicator_score_diff = indicator_score_std - 
                             lag(indicator_score_std, default = 0)) %>% 
                     mutate(indicator = ifelse(indicator == "RLI annual", "RLI",
                                               indicator)) %>% 
                    mutate(scenario_indicator = paste(scenario, indicator, 
                                                       sep = "_"),
                            variance = rollapply(indicator_score_diff,
                               10,
                               var,
                               na.rm = TRUE,
                               partial = TRUE,
                               align = "left"),
                            disturbance = as.integer(ifelse(annual_time_step > 99 &
                                                 annual_time_step < 201,
                                                 1, 0)),
                            disturbance_factor = as.factor(ifelse(disturbance == 1,
                                                 "disturbance", 
                                                 "no disturbance")),
                            years_since_disturbance = ifelse(annual_time_step < 100,
                                                            1, (annual_time_step - 100)),
                            disturbance_phase = as.integer(ifelse(annual_time_step < 100,
                                                                  1,
                                                 ifelse(annual_time_step > 99 &
                                                 annual_time_step < 201,
                                                 2, 3)))) %>%
                     ungroup(.)

# Split into a list of scenarios then indicators

scenario_indicators <- split(indicators_all_df, indicators_all_df$scenario)

# Or vice versa

indicator_scenarios <- split(indicators_all_df, indicators_all_df$indicator)

# Single indicator, single scenario

scenario_indicator_single <- split(indicators_all_df, indicators_all_df$scenario_indicator)

# Make a test df of one scenario, one indicator for building loop content

test_df <- indicators_all_df %>% 
           filter(indicator == "total abundance harvested") %>% 
           filter(scenario != "000_Baseline")

# Get the ylimits for indicator plots

# Specify y - limits of indicators

df <- indicators_all_df %>% 
      dplyr::filter(indicator  == "RLI") %>% 
      filter(scenario != "000_Baseline")

rlmin <- min(df$indicator_score) 
rlmax <- max(df$indicator_score)
rli_lims <- c(rlmin,rlmax)

df <- indicators_all_df %>% 
      dplyr::filter(indicator  == "LPI") %>% 
      filter(scenario != "000_Baseline")

lpmin <- min(df$indicator_score)
lpmax <- max(df$indicator_score)
lpi_lims <- c(lpmin,lpmax)



```

# Species gams

```{r species gams}

species_gams <- list()
species_gam_data <- list()
species_gam_plots <- list()

for (i in seq_along(scenario_redlist_data_sampled)) {
  
scenario_data <- scenario_redlist_data_sampled[[i]]

group_ids <- unique(scenario_data$group_id)

spp_mods <- list()
spp_data <- list()
spp_plots <- list()


  for (j in seq_along(group_ids)) {

  species_data <- scenario_data %>% 
                filter(group_id == group_ids[[j]]) %>% 
                mutate(std_abundance = standardize(ave_abundance)) %>% 
                ungroup(.)

  spp_mods[[j]] <- gam(std_abundance ~ s(annual_time_step, k = 10),
         select = TRUE,
         method = "GCV.Cp",
         data = species_data)

  spp_data[[j]] <- get_species_gam_data(species_data, spp_mods[[j]], 
                                              scenarios[[i]])

  spp_plots[[j]] <- plot_test_gam(spp_data[[j]])
  
  }

names(spp_plots) <- group_ids

df <- do.call(rbind, spp_data)

species_gam_data[[i]] <- df

species_gam_plots[[i]] <- spp_plots

}

```

``` {r influential species}

influential_species_plots <- list()
influential_species_data <- list()
all_responses <- list()
fgs <- list()

for (i in seq_along(species_gam_data)) {


temp <- species_gam_data[[i]] %>% 
        merge(groups[c("group_id", "nutrition_source", "harvested")], 
              by.x = "indicator",
              by.y = "group_id") %>% 
        #Get only the species where significant trends are detected
        filter(first_significant == TRUE) %>% 
        #Get only the species with significant trends at start or end of impact
        filter(annual_time_step >= 95 &
               annual_time_step <= 105|
               annual_time_step >= 195 &
               annual_time_step <= 205) %>% 
        dplyr::select(indicator, annual_time_step, 
                      predicted, first_derivative, nutrition_source) %>% 
        #Add a new variable that denotes whether we are looking at impact start or end event
        mutate(impact = ifelse(annual_time_step >= 95 &
               annual_time_step <= 105, "Impact start",
               ifelse(annual_time_step >= 195 &
               annual_time_step <= 205, "Impact stop", NA)))  
  
# Get start ----

# Look at response to impact start

temp2 <- temp %>% 
         filter(impact == "Impact start") %>% 
         dplyr::select(-annual_time_step) %>% # Remove time steps
         mutate(absolute_fd = abs(first_derivative)) # convert 1st deriv to absolute value
       
# idk what tf is going on here but arrange is not working as it should, so
# had to separate it from the rest of the pipe

temp3 <- temp2 %>% 
         arrange(desc(absolute_fd))

all_start <- temp3 %>% 
         group_by(indicator) %>% # group by spp
         filter(absolute_fd == max(absolute_fd)) # get the highest value per spp

if(scenarios_title[[i]] == "Carnivore harvesting") {

x <- all_start[1:n,] # Get n most influential spp

harv <- all_start %>% 
        filter(indicator == "11.67"|
                indicator == "11.68" )

start <-rbind(x, harv)

} else {

start <- all_start[1:n,] # Get n most influential spp

}

rm(temp2, temp3)

# Get stop ----

# Look at response to impact start

temp2 <- temp %>% 
         filter(impact == "Impact stop") %>% 
         dplyr::select(-annual_time_step) %>% # Remove time steps
         mutate(absolute_fd = abs(first_derivative)) # convert 1st deriv to absolute value
       
# idk what tf is going on here but arrange is not working as it should, so
# had to separate it from the rest of the pipe

temp3 <- temp2 %>% 
         arrange(desc(absolute_fd))

all_stop <- temp3 %>% 
         group_by(indicator) %>% # group by spp
         filter(absolute_fd == max(absolute_fd)) # get the highest value per spp

stop <- all_stop[1:n,] # Get n most influential spp

influential_species_data[[i]] <- rbind(start, stop) 

all_responses[[i]] <- rbind(all_start, all_stop)

# Add the harvested groups if they're not already there

if(scenarios_title[[i]] == "Carnivore harvesting") {
  
  hg <- c("11.67", "11.68", influential_species_data[[i]]$indicator)
  fgs[[i]] <- unique(hg)
  
} else if(scenarios_title[[i]] == "Herbivore harvesting") {
  
  hg <- c("10.67", "10.68", influential_species_data[[i]]$indicator)
  fgs[[i]] <- unique(hg)
  
} else {

fgs[[i]] <- unique(influential_species_data[[i]]$indicator)

}

influential_species_plots[[i]] <- species_gam_plots[[i]][fgs[[i]]] 

#walk(influential_species_plots[[i]], print)

}

```

```{r influential_species_plots}

influential_spp_plots <- list()

for (i in seq_along(species_gam_data)) {

fg <- influential_species_data[[i]]$indicator

data <- species_gam_data[[i]][species_gam_data[[i]]$indicator %in% fg,]

temp <- data %>% 
        merge(groups[c("group_id", "nutrition_source", "mass_lower", 
                       "harvested")], 
              by.x = "indicator",
              by.y = "group_id") %>% 
        filter(annual_time_step < 293) %>% 
        merge(influential_species_data[[i]][c("indicator", "impact")], 
              by = "indicator") %>% 
        mutate(lower_kg = mass_lower/1000,
               label = paste(lower_kg, "kg", nutrition_source, 
                             "(", impact, ")", indicator,  sep = " ")) %>% 
        arrange(impact,lower_kg)

influential_spp_plots[[i]] <- ggplot(temp) +
        geom_line(aes(annual_time_step, predicted, col = impact)) +
        geom_line(aes(x = annual_time_step, y = positive_trend,
                      col = impact),
                  col = "blue", size = 1) +
        geom_line(aes(x = annual_time_step, y = negative_trend,
                      col = nutrition_source),
                  col = "red", size = 1) +
        geom_point(aes(x = annual_time_step, y = observed, group = indicator,
                       col = impact),
                                        alpha = 0.1,
                   size = 1) +
        scale_color_manual(values = c(pos_col, neg_col)) +
        scale_fill_manual(values = c(pos_col, neg_col)) +
        geom_ribbon(aes(x = annual_time_step, 
                        ymin = lower_pred, 
                        ymax = upper_pred, 
                        group = indicator),
                                        alpha = 0.3) +
                  labs(x = "Time (years)",
                        y = "Standardized abundance") + 
        geom_vline(xintercept=c(100,200), linetype = "dotted") +
        facet_wrap( ~ label, ncol = 2) +
        plot_theme

rm(data, temp, fg)

}

walk(influential_spp_plots, print)

```

```{r select influential spp}

spp_plots <- list()

for (i in seq_along(influential_species_data)) {

title <-  scenarios_title[[i]]

if( title == "Baseline") {
  
start_fgs <- influential_species_data[[i]] %>% 
     filter(impact == "Impact start") %>% 
     dplyr::select(indicator) %>% 
     pull(.)
start_fgs <- start_fgs[1:3]

stop_fgs <- influential_species_data[[i]] %>% 
     filter(impact == "Impact stop") %>% 
     dplyr::select(indicator) %>% 
     pull(.)

stop_fgs <- stop_fgs[1:3]

} else if (title == "Land use") {
  
  start_fgs <- c("10.33","15.18.39", "15.18.53")
  stop_fgs <- c("10.47","14.17.54", "10.72")
  
} else if (title == "Carnivore harvesting") {
  
  start_fgs <- c("11.67","12.66", "10.72")
  stop_fgs <- c("15.18.25","13.16.53", "12.65")
  
} else if (title == "Herbivore harvesting") {
  
  start_fgs <- c("10.68","10.72", "10.76")
  stop_fgs <- c("14.17.38","12.73", "13.16.57")
  
} 

data <- species_gam_data[[i]][species_gam_data[[i]]$indicator %in% start_fgs,]

temp <- data %>% 
        merge(groups[c("group_id", "nutrition_source", "mass_lower", 
                       "harvested")], 
              by.x = "indicator",
              by.y = "group_id") %>% 
    # merge(example_species[c("group_id", "common_name")],
    #                                    by.x = "indicator",
    #            by.y = "group_id") %>% 
        filter(annual_time_step < 293) %>% 
        merge(influential_species_data[[i]][c("indicator", "impact")], 
              by = "indicator") %>% 
        filter(impact == "Impact start") %>%
        mutate(lower_kg = mass_lower/1000,
               label = paste(lower_kg, "kg", nutrition_source, 
                             "(", indicator, ")", harvested, sep = " "),
               impact = "Impact start") %>% 
        arrange(impact,lower_kg)



start_plot <- ggplot(temp) +
        geom_line(aes(annual_time_step, predicted)) +
        geom_line(aes(x = annual_time_step, y = positive_trend),
                  col = pos_col, size = 2) +
        geom_line(aes(x = annual_time_step, y = negative_trend,
                      col = nutrition_source),
                  col = neg_col, size = 2) +
        geom_point(aes(x = annual_time_step, y = observed, group = indicator),
                                        alpha = 0.1,
                   size = 0.5) +
        scale_color_manual(values = c(pos_col, neg_col)) +
        scale_fill_manual(values = c(pos_col, neg_col)) +
        geom_ribbon(aes(x = annual_time_step, 
                        ymin = lower_pred, 
                        ymax = upper_pred, 
                        group = indicator),
                                        alpha = 0.3) +
                  labs(x = "Time (years)",
                        y = "Standardized abundance") + 
        geom_vline(xintercept=c(100,200), linetype = "dotted") +
        facet_wrap(~ label, nrow = 3) +
        theme(panel.background = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              legend.position = "bottom",
              legend.key.size = unit(0.25, 'cm'),
              plot.title = element_text(size = 10, face = "bold"),
              legend.title=element_text(size=6), 
              legend.text=element_text(size=5),
              axis.line = element_line(colour = "black"),
              axis.title.x = element_text(size=8),
              axis.title.y = element_text(size=8),
              axis.text.x = element_text(size=8),
              axis.text.y = element_text(size=8),
              axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank(),
              strip.text.x = element_text(size = 6))

data <- species_gam_data[[i]][species_gam_data[[i]]$indicator %in% stop_fgs,]

temp <- data %>% 
        merge(groups[c("group_id", "nutrition_source", "mass_lower", 
                       "harvested")], 
              by.x = "indicator",
              by.y = "group_id") %>% 
  # merge(example_species[c("group_id", "common_name")],
  #                                      by.x = "indicator",
  #              by.y = "group_id") %>% 
        filter(annual_time_step < 293) %>% 
        merge(influential_species_data[[i]][c("indicator", "impact")], 
              by = "indicator") %>% 
        filter(impact == "Impact stop") %>% 
        mutate(lower_kg = mass_lower/1000,
               label = paste(lower_kg, "kg", nutrition_source, 
                             "(", indicator, ")",harvested, sep = " ")) %>% 
        arrange(impact,lower_kg)

stop_plot <- ggplot(temp) +
        geom_line(aes(annual_time_step, predicted)) +
        geom_line(aes(x = annual_time_step, y = positive_trend),
                  col = pos_col, size = 2) +
        geom_line(aes(x = annual_time_step, y = negative_trend,
                      col = nutrition_source),
                  col = neg_col, size = 2) +
        geom_point(aes(x = annual_time_step, y = observed, group = indicator),
                   alpha = 0.1, size = 0.5) +
        scale_color_manual(values = c(pos_col, neg_col)) +
        scale_fill_manual(values = c(pos_col, neg_col)) +
        geom_ribbon(aes(x = annual_time_step, 
                        ymin = lower_pred, 
                        ymax = upper_pred, 
                        group = indicator),
                                        alpha = 0.3) +
                  labs(x = "Time (years)",
                        y = "Standardized abundance") + 
        geom_vline(xintercept=c(100,200), linetype = "dotted") +
        facet_wrap(~ label, nrow = 3) +
        theme(panel.background = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              legend.position = "bottom",
              legend.key.size = unit(0.25, 'cm'),
              plot.title = element_text(size = 10, face = "bold"),
              legend.title=element_text(size=6), 
              legend.text=element_text(size=5),
              axis.line = element_line(colour = "black"),
              axis.title.x = element_text(size=8),
              axis.title.y = element_text(size=8),
              axis.text.x = element_text(size=8),
              axis.text.y = element_text(size=8),
              axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank(),
              strip.text.x = element_text(size = 6))

spp_plots[[i]] <- plot_grid(start_plot, stop_plot, ncol = 2, align = "v")

ggsave(file.path(manuscript_plots_folder, paste(scenarios[[i]],
                                                "species_response_gams.png",
                                                sep = "_")),
       width = 14,
       height = 8,
       units = "cm",
       spp_plots[[i]])

rm(data, temp, stop_fgs, start_plot, stop_plot)

}

walk(spp_plots, print)

```

# All responses

```{r response plots}

response_boxplots <- list()

for (i in seq_along(all_responses)) {
  
bxplot_data <- all_responses[[i]] %>% 
               merge(groups[c("group_id", "functional_group_name", "mass_lower")], 
              by.x = "indicator",
              by.y = "group_id") %>% 
              mutate(mass_lower_kg = mass_lower/1000)

response_boxplots[[i]] <- ggplot(bxplot_data, aes(x = functional_group_name, 
                                    y = first_derivative)) +
                            geom_violin(alpha = 0.1, aes(col = functional_group_name)) +
            geom_point(aes(x = functional_group_name,
                           y = first_derivative,
                           fill = functional_group_name,
                           col = functional_group_name,
                           size = mass_lower_kg), alpha = 0.4) +
            #geom_jitter(aes(alpha = mass_lower_kg)) +
            geom_hline(yintercept = 0, linetype = "dotted") +
            labs(y = "Functional species response") +
            theme(panel.background = element_rect(fill = 'grey83'),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              legend.position = "bottom",
              legend.key.size = unit(0.25, 'cm'),
              plot.title = element_text(size = 10, face = "bold"),
              legend.title=element_text(size=7), 
              legend.text=element_text(size=6),
              axis.line = element_line(colour = "black"),
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              axis.text.x = element_text(size=8),
              axis.text.y = element_text(size=8),
              axis.ticks.x = element_blank(), 
              axis.ticks.y = element_blank()) +
             facet_wrap(~impact) +
  scale_fill_viridis_d(guide = "none", option = "mako") +
  scale_color_viridis_d(guide = "none",option = "mako") +
  #scale_colour_hue(guide = "none") +
  scale_fill_hue(guide = "none")+ 
  scale_y_continuous(breaks = c(-0.08, 0, 0.08),
                     labels = c("Decreased \n in abundance",
                                "No change",
                                "Increased \n in abundance"),
                     limits = c(-0.1,0.1)) +
  labs(size = "Body size (kg)") +
  coord_flip()

ggsave(file.path(manuscript_plots_folder, paste(scenarios[[i]],
                                                "functional_group_boxplots.png",
                                                sep = "_")),
       width = 14,
       height = 8,
       units = "cm",
       response_boxplots[[i]])


}

walk(response_boxplots, print)

```


# Functional group gams

```{r fg gams}

fg_gams <- list()
fg_gam_data <- list()
fg_gam_plots <- list()

for (i in seq_along(scenario_redlist_data_sampled)) {
  
scenario_data <- scenario_redlist_data_sampled[[i]] %>% 
                 group_by(annual_time_step,functional_group_name) %>% 
                 mutate(fg_abundance = sum(ave_abundance, na.rm = TRUE)) %>% 
                 dplyr::select(annual_time_step, fg_abundance, functional_group_name) %>% 
                 distinct(.) %>% 
                 ungroup(.) %>% 
                 group_by(functional_group_name) %>% 
                 mutate(std_abundance = standardize(fg_abundance)) %>% 
                 ungroup(.)

fg_ids <- unique(scenario_data$functional_group_name)

fg_mods <- list()
fg_data <- list()
fg_plots <- list()


  for (j in seq_along(fg_ids)) {

  data <- scenario_data %>% 
          filter(functional_group_name == fg_ids[[j]]) %>% 
          rename(group_id = functional_group_name)
  
  fg_mods[[j]] <- gam(std_abundance ~ s(annual_time_step, k = 10),
         select = TRUE,
         method = "GCV.Cp",
         data = data)
  
  fg_data[[j]] <- get_species_gam_data(data, fg_mods[[j]], 
                                              scenarios[[i]])
  
  #fg_plots[[j]] <- plot_test_gam(fg_data[[j]])
  
  
  
  }

# names(fg_data) <- fg_ids
# names(fg_mods) <- fg_ids

df <- do.call(rbind, fg_data) %>% 
      rename(functional_group = indicator)

fg_gam_plots[[i]] <- ggplot(df) +
        geom_line(aes(annual_time_step, predicted)) +
        geom_line(aes(x = annual_time_step, y = positive_trend), 
                  col = "blue", size = 1) +
        geom_line(aes(x = annual_time_step, y = negative_trend), 
                  col = "red", size = 1) +
        geom_point(aes(x = annual_time_step, y = observed),
                                        alpha = 0.1,
                   size = 1) +
        geom_ribbon(aes(x = annual_time_step, 
                        ymin = lower_pred, 
                        ymax = upper_pred),
                                        alpha = 0.3) +
                  labs(x = "Time (years)",
                        y = "Standardized abundance",
                       title = scenarios[[i]]) + 
        geom_vline(xintercept=c(100,200), linetype = "dotted") +
        facet_wrap( ~ functional_group) +
        plot_theme

ggsave(file.path(manuscript_plots_folder, paste(scenarios[[i]],
                                                "functional_group_gams.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
       fg_gam_plots[[i]])

fg_gam_data[[i]] <- df

}

walk(fg_gam_plots, print)

```

# Trophic group gams

``` {r tg gams}

tg_gams <- list()
tg_gam_data <- list()
tg_gam_plots <- list()

for (i in seq_along(scenario_redlist_data_sampled)) {

auto <- auto_annual[[i]] %>% 
        filter(annual_time_step > 1) %>% 
        filter(annual_time_step < 293) %>% 
        mutate(functional_group_name = group_id) %>% 
        filter(group_id == "autotrophs")

all <- rbind(scenario_redlist_data_sampled[[i]], auto)

scenario_data <- all  %>% 
                 mutate(trophic_group = word(functional_group_name, 1)) %>% 
                 group_by(annual_time_step,trophic_group) %>% 
                 mutate(tg_abundance = sum(ave_abundance, na.rm = TRUE)) %>% 
                 dplyr::select(annual_time_step, tg_abundance, trophic_group) %>% 
                 distinct(.) %>% 
                 ungroup(.) %>% 
                 group_by(trophic_group) %>% 
                 mutate(std_abundance = standardize(tg_abundance)) %>% 
                 ungroup(.)

tg_ids <- unique(scenario_data$trophic_group)

tg_mods <- list()
tg_data <- list()
tg_plots <- list()

  for (j in seq_along(tg_ids)) {

  data <- scenario_data %>% 
          filter(trophic_group == tg_ids[[j]]) %>% 
          rename(group_id = trophic_group)
  
  tg_mods[[j]] <- gam(std_abundance ~ s(annual_time_step, k = 10),
         select = TRUE,
         method = "GCV.Cp",
         data = data)
  
  tg_data[[j]] <- get_species_gam_data(data, tg_mods[[j]], 
                                              scenarios[[i]])
  
 
  }

df <- do.call(rbind, tg_data) %>% 
      rename(trophic_group = indicator)

tg_gam_plots[[i]] <- ggplot(df) +
        geom_line(aes(annual_time_step, predicted)) +
        geom_line(aes(x = annual_time_step, y = positive_trend), 
                  col = "blue", size = 1) +
        geom_line(aes(x = annual_time_step, y = negative_trend), 
                  col = "red", size = 1) +
        geom_point(aes(x = annual_time_step, y = observed),
                                        alpha = 0.1,
                   size = 1) +
        geom_ribbon(aes(x = annual_time_step, 
                        ymin = lower_pred, 
                        ymax = upper_pred),
                                        alpha = 0.3) +
                  labs(x = "Time (years)",
                        y = "Standardized abundance",
                       title = scenarios[[i]]) + 
        geom_vline(xintercept=c(100,200), linetype = "dotted") +
        facet_wrap( ~ trophic_group) +
        plot_theme

ggsave(file.path(manuscript_plots_folder, paste(scenarios[[i]],
                                                "trophic_group_gams.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
       tg_gam_plots[[i]])

tg_gam_data[[i]] <- df

}

walk(tg_gam_plots, print)

```



# All test gams

```{r all test gams, warnings = FALSE, message = FALSE }

ind_method <- "GCV.Cp"
selected <- 1

all_mods_data <- list()
selected_mods <- list()
selected_mods_data <- list()
robust_trends <- list()
deviance_explained <- list()
plots <- list()

for ( i in seq_along(scenario_indicator_single)) {

#scenario <- scenario_indicator_single[[i]]$scenario[1]
scenario <- scenarios_title_2[[i]]
indicator <- scenario_indicator_single[[i]]$indicator[1]

indicator <- str_replace(indicator, "/", "_")

kvals <- c(10,25,50)

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
         select = TRUE,
         method = ind_method,
         data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2)

# Extract observed, predicted, derivatives and plot for each test gam

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
# plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]], scenario)

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

all_mods_data[[i]] <- test_gam_data

plots[[i]] <- plot_grid(test_gam_plots[[1]],
                        test_gam_plots[[2]],
                        test_gam_plots[[3]],
                        nrow = 2, align = "v")

print(plots[[i]])

ggsave(file.path(supporting_info_plots_folder, paste(scenario, indicator,
                                                "test_gams.png",
                                                sep = "_")),
       width = t_width,
       height = t_height,
       units = "cm",
       plots[[i]])

robust_trends[[i]] <- compare_trends(test_gam_data)

selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

}

```

# All bootstrapped GAMS

```{r model baseline lpi, warnings = FALSE }

rm(bootstrapped_plots)

bootstrapped_data <- list()
bootstrapped_plots <- list()

for (i in seq_along(selected_mods)) {
  
scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]
indicator <- str_replace(indicator, "/", "_")

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 3,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models == 3,
                                                predicted,NA))

if (indicator == "RLI") {
  
  lims <- c(0.85, 1.00)
  
} else if (indicator == "LPI") {
  
  lims <- c(0.75, 1.15)
}
# hamin <- min(bootstrapped_data[[i]]$predicted_upper) - 0.2
# hamax <- max(bootstrapped_data[[i]]$predicted_lower) + 0.2
# lims <- c(hamin,hamax)


bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   lims)

print("plot function")
print(i)

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

print("save csv")
print(i)

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bt.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
       bootstrapped_plots[[i]])


bootstrapped_plots[[i]] <- bootstrapped_plots[[i]] +
                            labs(title = scenario)

print("add title")
print(i)


}

walk(bootstrapped_plots, print)

```

# Trends

```{r summarise trends}

trends <- list()

for (i in seq_along(all_mods_data)) {

  df <- do.call(rbind, all_mods_data[[i]])
 
  scen <- scenarios_title_2[i]
  scen
  # Get the years at which a robust trend is observed
  
  yrs <- robust_trends[[i]] %>% 
       mutate(robust = ifelse(negative_across_all_models == 3,
                              TRUE,
              ifelse(positive_across_all_models == 3, TRUE, FALSE)),
              direction = ifelse(negative_across_all_models == 3,
                              "negative",
              ifelse(positive_across_all_models == 3, "positive", NA))) %>% 
       dplyr::select(annual_time_step, robust, direction) %>% 
          mutate(impact = ifelse(annual_time_step >= 95 &
               annual_time_step <= 105, "Impact start",
               ifelse(annual_time_step >= 195 &
               annual_time_step <= 205, "Impact stop", NA))) %>% 
      filter(!is.na(impact)) %>% 
      dplyr::select(-impact)
  
  
  trend <- df %>% 
          mutate(impact = ifelse(annual_time_step >= 95 &
               annual_time_step <= 105, "Impact start",
               ifelse(annual_time_step >= 195 &
               annual_time_step <= 205, "Impact stop", NA))) %>% 
          filter(impact =="Impact start"|
                 impact == "Impact stop") %>%
          filter(kvalue == 10) %>% 
          dplyr::select(annual_time_step, indicator, scenario, predicted,
                        first_derivative, first_significant, impact) %>% 
          merge(yrs, by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_trend = ifelse(is.na(robust), NA ,
                                       first_derivative),
                 scenario2 = scen) %>% 
          group_by(impact) %>% 
          mutate(ave_trend = mean(robust_trend, na.rm = TRUE)) %>% #Get the average trend 
          dplyr::select(impact, indicator, scenario, scenario2, ave_trend) %>% 
          distinct(.) %>% 
          mutate(ave_trend = ifelse(is.nan(ave_trend), 0, ave_trend))
  
  scenario_fac <- factor(trend$scenario2,
      levels = c("Baseline", "Harvesting primary producers",
                 "Harvesting herbivores", "Harvesting carnivores"))

 trends[[i]] <- cbind(trend,
  scenario_fac = scenario_fac)
  
}

# See all trends
all_robust <- list()

for (i in seq_along(robust_trends)) {
  
  all_robust[[i]] <- robust_trends[[i]] %>% 
                     mutate(id = names(scenario_indicator_single[i]))
}

all_robust_df <- do.call(rbind, all_robust)

```

# Trends

```{r plot trends}

trend_table <- do.call(rbind, trends) %>% 
               mutate(ave_trend = round(ave_trend, 3)) %>% 
               filter(scenario != "Baseline")

trend_table_wide <- trend_table %>%
                dplyr::select(scenario, indicator, impact, ave_trend) %>% 
               pivot_wider(names_from = c(scenario, impact),
                           values_from = ave_trend)

write.csv(trend_table_wide, file.path(manuscript_plots_folder, 
                   "trend_table.csv"))

plot_data <- trend_table %>% 
            mutate(label = paste(scenario, impact))


heatmap <- ggplot(plot_data, aes(x = impact, y = indicator, 
                                 fill = ave_trend)) +
  geom_tile(color = "white", height = 1.0, width = 1.0, alpha = 0.75) +
  coord_equal() +
  geom_text(aes(label = round(ave_trend, 3)), size = 2) +
  scale_fill_gradient2(
    low = neg_col, mid = 'white', high = pos_col,
    midpoint = 0, guide = 'colourbar', aesthetics = 'fill'
  ) +
  theme(axis.text.x = element_text(size=8, angle = 45, vjust = 0.5, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none") +
  facet_grid(~ scenario_fac) 

heatmap

ggsave(file.path(manuscript_plots_folder, "results_heatmap.png"),
       width = 18,
       height = 10,
       units = "cm",
       heatmap)


```
# Baseline test gams
## LPI

First derivatives show the rate of change in the indicator at a specific time
point. Values = 0 signify no trend, values > 0 signify a positive trend, and
values < 0 signify a negative trend.

"The second derivative of the trend curve I(t) at time t is a measure of the curvature of I at that time. If the second derivative is greater than zero, the curve is turning upward. If it is less than zero, the curve is turning downward. The magnitude of the second derivative signifies the tightness of the curvature. Values of approx 0 indicate the curve is linear and changing at a steady state. Years in which the second derivative is markedly different from zero are those in which something is happening to alter the rate of population change." (Fewster et al 2000)

```{r baseline lpi, warnings = FALSE, message = FALSE }

selected_mods <- list()
selected_mods_data <- list()
robust_trends <- list()
deviance_explained <- list()

i <- 1

scenario <- "Baseline"
indicator <- scenario_indicator_single[[i]]$indicator[1]

kvals <- c(10,25,50)

ind_method <- "GCV.Cp"

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
         select = TRUE,
         method = ind_method,
         data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2)

# Extract observed, predicted, derivatives and plot for each test gam

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}



plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 nrow = 2, align = "v")

plots

ggsave(file.path(supporting_info_plots_folder, paste(scenario, indicator,
                                                "test_gams.png",
                                                sep = "_")),
       width = t_width,
       height = t_height,
       units = "cm",
       plots)

robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## RLI

```{r baseline rli, message=FALSE, warnings=FALSE}

i <- 2

indicator <- scenario_indicator_single[[i]]$indicator[1]

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
         select = TRUE,
         method = ind_method,
         data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

plots

ggsave(file.path(supporting_info_plots_folder, paste(scenario, indicator,
                                                "test_gams.png",
                                                sep = "_")),
       width = t_width,
       height = t_height,
       units = "cm",
       plots)

robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## Harvested group

```{r baseline harvesed, warnings = FALSE, message = FALSE}

i <- 3

indicator <- scenario_indicator_single[[i]]$indicator[1]

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

plots

ggsave(file.path(supporting_info_plots_folder, paste(scenario, indicator,
                                                "test_gams.png",
                                                sep = "_")),
       width = t_width,
       height = t_height,
       units = "cm",
       plots)



robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

# Land use test gams
## LPI

```{r land use lpi, warnings = FALSE, message = FALSE }

# Make first three items NULL (first 3 elements are taken up by baseline data 
# which we aren't analysing here)

# robust_trends[[1]] <- NULL
# robust_trends[[2]] <- NULL
# robust_trends[[3]] <- NULL
# 
# deviance_explained[[1]] <- NULL
# deviance_explained[[2]] <- NULL
# deviance_explained[[3]] <- NULL

i <- 4 

indicator <- scenario_indicator_single[[i]]$indicator[1]

scenario <- "Land use"

kvals <- c(10,25,50,100)

ind_method <- "GCV.Cp"

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
         select = TRUE,
         method = ind_method,
         data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

# Extract observed, predicted, derivatives and plot for each test gam

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

residual_plots

plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

plots

ggsave(file.path(supporting_info_plots_folder, paste(scenario, indicator,
                                                "test_gams.png",
                                                sep = "_")),
       width = t_width,
       height = t_height,
       units = "cm",
       plots)

robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## RLI

```{r land use rli, message=FALSE, warnings=FALSE}

i <- 5

indicator <- scenario_indicator_single[[i]]$indicator[1]

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
         select = TRUE,
         method = ind_method,
         data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

plots

ggsave(file.path(supporting_info_plots_folder, paste(scenario, indicator,
                                                "test_gams.png",
                                                sep = "_")),
       width = t_width,
       height = t_height,
       units = "cm",
       plots)


robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## Harvested group

```{r harvested rli, warnings = FALSE, message = FALSE}

i <- 6

indicator <- scenario_indicator_single[[i]]$indicator[1]

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

plots

ggsave(file.path(supporting_info_plots_folder, paste(scenario, indicator,
                                                "test_gams.png",
                                                sep = "_")),
       width = t_width,
       height = t_height,
       units = "cm",
       plots)

robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```


# Harvesting carnivores test gams
## LPI

```{r carnivores lpi, warnings = FALSE, message = FALSE }

i <- 7

scenario <- "Harvesting carnivores"

indicator <- scenario_indicator_single[[i]]$indicator[1]

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

plots

ggsave(file.path(supporting_info_plots_folder, paste(scenario, indicator,
                                                "test_gams.png",
                                                sep = "_")),
       width = t_width,
       height = t_height,
       units = "cm",
       plots)

robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## RLI

```{r carnivores rli, warnings = FALSE, message = FALSE }

i <- 8 
indicator <- scenario_indicator_single[[i]]$indicator[1]

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

plots

ggsave(file.path(supporting_info_plots_folder, paste(scenario, indicator,
                                                "test_gams.png",
                                                sep = "_")),
       width = t_width,
       height = t_height,
       units = "cm",
       plots)

robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## Harvested group

```{r carnivores harvested, warnings = FALSE, message = FALSE }

i <- 9
indicator <- scenario_indicator_single[[i]]$indicator[1]

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

plots

ggsave(file.path(supporting_info_plots_folder, paste(scenario, indicator,
                                                "test_gams.png",
                                                sep = "_")),
       width = t_width,
       height = t_height,
       units = "cm",
       plots)


robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

# Harvesting herbivores test gams
## LPI

```{r herbivores lpi, warnings = FALSE }

i <- 10

scenario <- "Harvesting herbivores"
indicator <- scenario_indicator_single[[i]]$indicator[1]

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

plots

ggsave(file.path(supporting_info_plots_folder, paste(scenario, indicator,
                                                "test_gams.png",
                                                sep = "_")),
       width = t_width,
       height = t_height,
       units = "cm",
       plots)

robust_trends[[i]] <- compare_trends(test_gam_data)

trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)

selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## RLI

```{r herbivores rli, warnings = FALSE }

i <- 11
indicator <- scenario_indicator_single[[i]]$indicator[1]

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

residual_plots

plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

plots

ggsave(file.path(supporting_info_plots_folder, paste(scenario, indicator,
                                                "test_gams.png",
                                                sep = "_")),
       width = t_width,
       height = t_height,
       units = "cm",
       plots)

robust_trends[[i]] <- compare_trends(test_gam_data)


trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)


selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

## Harvested group

```{r herbivores harvested, warnings = FALSE }

i <- 12 
indicator <- scenario_indicator_single[[i]]$indicator[1]

m <- gam(indicator_score ~ s(annual_time_step, k = kvals[1]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m1 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[2]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m2 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[3]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

m3 <- gam(indicator_score ~ s(annual_time_step, k = kvals[[4]]),
          select = TRUE,
          method = ind_method,
          data = scenario_indicator_single[[i]])

indicator_mods <- list(m, m1, m2, m3)

test_gam_data <- list()
test_gam_plots <- list()
residual_plots <- list()

for (j in seq_along(indicator_mods)) {
  
plot.gam(indicator_mods[[j]], residuals = TRUE)

test_gam_data[[j]] <- get_test_gam_data(scenario_indicator_single[[i]],
                               indicator_mods[[j]])

test_gam_plots[[j]] <- plot_test_gam(test_gam_data[[j]])

}

residual_plots

plots <- plot_grid(test_gam_plots[[1]],
                                 test_gam_plots[[2]],
                                 test_gam_plots[[3]],
                                 test_gam_plots[[4]],
                                 nrow = 2, align = "v")

plots

ggsave(file.path(supporting_info_plots_folder, paste(scenario, indicator,
                                                "test_gams.png",
                                                sep = "_")),
       width = t_width,
       height = t_height,
       units = "cm",
       plots)

trend_years <- robust_trends[[i]] %>% 
          filter(positive_across_all_models == 1|
                   negative_across_all_models == 1)


selected <- 1
selected_mods[[i]] <- indicator_mods[[selected]]
selected_mods_data[[i]] <- test_gam_data[[selected]]

deviance_explained[[i]] <- round(summary.gam(selected_mods[[i]])$dev.expl, 2)

```

# Baseline selected model
## LPI

```{r model baseline lpi, warnings = FALSE }

bootstrapped_data <- list()
bootstrapped_plots <- list()

i <- 1

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]], 
                                                   lpi_lims)


scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
       bootstrapped_plots[[i]])

bootstrapped_plots[[i]]

```

## RLI

```{r model baseline rli, warnings = FALSE }

i <- 2

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots)%>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   rli_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]


```

## Harvested group

```{r model baseline harvested, warnings = FALSE }

i <- 3

df <- indicators_all_df %>% 
      dplyr::filter(indicator  == "total abundance harvested") %>% 
      filter(scenario == "000_Baseline")

hmin <- min(df$indicator_score)
hmax <- max(df$indicator_score)

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   c(hmin, hmax))

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))


ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]



```

# Land use selected model
## LPI

```{r model land use lpi, warnings = FALSE }

bootstrapped_data <- list()
bootstrapped_plots <- list()

i <- 4

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]], 
                                                   lpi_lims)


scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
       bootstrapped_plots[[i]])

bootstrapped_plots[[i]]

```

## RLI

```{r model land use rli, warnings = FALSE }

i <- 5

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots)%>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   rli_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]


```

## Harvested group

```{r model land use harvested, warnings = FALSE }

i <- 6

# hmax <- get_max(scenario_indicator_single[[i]], "total_abundance_harvested")
# hmin <- get_min(scenario_indicator_single[[i]], "total_abundance_harvested")

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   harv_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))


ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]



```

# Harvesting carnivores selected model
## LPI

```{r model carnivores lpi, warnings = FALSE }

i <- 7

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   lpi_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]


```

## RLI

```{r model carnivores rli, warnings = FALSE }

i <- 8 

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   rli_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]

```

## Harvested group

```{r model carnivores harvested, warnings = FALSE }

i <- 9

df <- indicators_all_df %>% 
      dplyr::filter(indicator  == "total abundance harvested") %>% 
      filter(scenario == "200_Harvesting_carnivores")

hmin <- min(df$indicator_score)
hmax <- max(df$indicator_score)


bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   c(hmin,hmax))

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]


```

# Harvesting herbivores selected models
## LPI

```{r model herbivores lpi, warnings = FALSE }

i <- 10

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   lpi_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))


ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]

```

## RLI

```{r model herbivores rli, warnings = FALSE }

i <- 11

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   rli_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]

```

## Harvested group

```{r model herbivores harvested, warnings = FALSE }

i <- 12 

# hmax <- get_max(scenario_indicator_single[[i]], "total_abundance_harvested")
# hmin <- get_min(scenario_indicator_single[[i]], "total_abundance_harvested")

bootstrapped_data[[i]] <- bootstrap_selected_model(scenario_indicator_single[[i]],
                                 selected_mods[[i]], 
                                 selected_mods_data[[i]], boots) %>% 
          merge(robust_trends[[i]], by = "annual_time_step", all = TRUE) %>% 
          mutate(robust_positive_trend = ifelse(positive_across_all_models == 1,
                                                predicted,NA)) %>% 
          mutate(robust_negative_trend = ifelse(negative_across_all_models== 1,
                                                predicted,NA))

bootstrapped_plots[[i]] <- plot_bootstrapped_model(bootstrapped_data[[i]],
                                                   harv_lims)

scenario <- scenario_indicator_single[[i]]$scenario[1]
indicator <- scenario_indicator_single[[i]]$indicator[1]

write.csv(bootstrapped_data[[i]], file.path(analysis_outputs_folder,
                                            paste(scenario, indicator,
                                                "bootstrapped_trends.csv",
                                                sep = "_")))

ggsave(file.path(manuscript_plots_folder, paste(scenario, indicator,
                                                "bootstrapped_trend.png",
                                                sep = "_")),
       width = 10,
       height = 5,
       units = "cm",
                 bootstrapped_plots[[i]])

bootstrapped_plots[[i]]


```

# Get periods of change

```{r periods of change, results=FALSE }

all_boot_data <- do.call(rbind, bootstrapped_data) %>% 
                 dplyr::select(annual_time_step, scenario, indicator,
                               robust_positive_trend, robust_negative_trend) %>%
                 filter_at(vars(robust_positive_trend,
                                robust_negative_trend),any_vars(!is.na(.)))
test <- all_boot_data %>% 
        filter(scenario == "100_Land_use") %>% 
        pivot_wider(names_from = indicator,
                    values_from = robust_positive_trend)



```
<!-- # Using MGCV All models -->

<!-- ```{r m all, echo = TRUE} -->

<!-- selected_mods <- list() -->

<!-- for(i in seq_along(scenario_indicator_single)) { -->

<!-- print(scenario_indicator_single[[i]]$scenario_indicator[1]) -->

<!-- selected_mods[[i]] <- gamm(indicator_score ~ s(annual_time_step), -->
<!--           select = TRUE, -->
<!--           data = scenario_indicator_single[[i]], -->
<!--           correlation = corCAR1(form = ~ 1 | annual_time_step)) -->

<!-- } -->

<!-- names(selected_mods) <- names(scenario_indicator_single) -->

<!-- gam.check(selected_mods[[1]]$gam) -->
<!-- names(selected_mods)[1] #LPI -->
<!-- gam.check(selected_mods[[2]]$gam) -->
<!-- names(selected_mods)[2] # RLI -->
<!-- gam.check(selected_mods[[3]]$gam)  -->
<!-- names(selected_mods)[3] # baseline abundance, doesn't matter if weird  -->
<!-- gam.check(selected_mods[[4]]$gam) -->
<!-- names(selected_mods)[4] #LPI -->
<!-- gam.check(selected_mods[[5]]$gam) -->
<!-- names(selected_mods)[5] #RLI -->
<!-- gam.check(selected_mods[[6]]$gam)  -->
<!-- names(selected_mods)[6] # total abundance harvested, may not matter either -->
<!-- gam.check(selected_mods[[7]]$gam) -->
<!-- names(selected_mods)[7] #LPI -->
<!-- gam.check(selected_mods[[8]]$gam)  -->
<!-- names(selected_mods)[8] # RLI carnivores, IMPORTANT, needs more work -->
<!-- gam.check(selected_mods[[9]]$gam)  -->
<!-- names(selected_mods)[9] # Abundance carnivores, may not matter -->
<!-- gam.check(selected_mods[[10]]$gam)  -->
<!-- names(selected_mods)[10] # LPI herbivores, IMPORTANT, needs more work -->
<!-- gam.check(selected_mods[[11]]$gam) -->
<!-- names(selected_mods)[11] # RLI -->
<!-- gam.check(selected_mods[[12]]$gam)  -->
<!-- names(selected_mods)[12] # Abundance harvested, may not matter -->

<!-- ``` -->




<!-- # Plot fitted against observed -->

<!-- ```{r model selection 1, echo = TRUE } -->

<!-- fitted_observed_plots <- list() -->

<!-- for (i in seq_along(scenario_indicator_single)) { -->

<!-- fit_obs_df <- data.frame(annual_time_step = scenario_indicator_single[[i]]$annual_time_step,  -->
<!--                    fitted = selected_mods[[i]]$gam$fitted.values,  -->
<!--                    observed = scenario_indicator_single[[i]]$indicator_score) -->

<!-- fitted_observed_plots[[i]] <- ggplot() +  -->
<!--   geom_line(data = fit_obs_df, aes(annual_time_step, fitted)) +  -->
<!--   geom_point(data = fit_obs_df, aes(annual_time_step, observed)) -->

<!-- } -->

<!-- fitted_observed_plots[[1]] -->
<!-- plot_trend(selected_mods2[[1]]) -->
<!-- fitted_observed_plots[[2]] -->
<!-- fitted_observed_plots[[3]] -->
<!-- fitted_observed_plots[[4]] -->
<!-- fitted_observed_plots[[5]] -->
<!-- fitted_observed_plots[[6]] -->
<!-- fitted_observed_plots[[7]] -->
<!-- fitted_observed_plots[[8]] -->
<!-- fitted_observed_plots[[9]] -->
<!-- fitted_observed_plots[[10]] -->
<!-- fitted_observed_plots[[11]] -->
<!-- fitted_observed_plots[[12]] -->

<!-- ``` -->

<!-- # Graveyard  -->
<!-- ```{r graveyard, echo = TRUE } -->

<!-- # Using Jess' method -->

<!-- test_plots <- list() -->

<!-- d <- 2 -->

<!-- for (i in seq_along(scenario_indicator_single)) { -->

<!-- fit_obs_df <- data.frame(annual_time_step = scenario_indicator_single[[i]]$annual_time_step,  -->
<!--                    fitted = selected_mods[[i]]$gam$fitted.values,  -->
<!--                    observed = scenario_indicator_single[[i]]$indicator_score_std) -->

<!-- deriv_func <- splinefun(x = fit_obs_df$annual_time_step, y = fit_obs_df$fitted) -->
<!-- deriv_fitted <- as.data.frame(x = deriv_func(seq(1, 290, 1), deriv = d)) -->

<!-- deriv_df <- cbind(ts = fit_obs_df$annual_time_step,d = deriv_fitted, -->
<!--                   observed = fit_obs_df$observed) -->

<!-- names(deriv_df) <- c("annual_time_step", "deriv", "observed") -->

<!-- test_plots[[i]] <- ggplot() + -->
<!--                    #geom_point(data = deriv_df,aes(annual_time_step, observed)) + -->
<!--                    geom_line(data = deriv_df,aes(annual_time_step, deriv)) -->



<!-- } -->

<!-- i <- i + 1 -->
<!-- test_plots[[i]] -->





<!-- ``` -->

<!-- # 2nd derivatives -->

<!-- Second derivatives show when the indicator values have changed direction. -->
<!-- "The second derivative of the trend curbe I(t) at time t is a measure of the curvature of I at that time. If the second derivative is greater than zero, the curve is turning upward. If it is less than zero, the curve is turning downward. The magnitude of the second derivative signifies the tightness of the curvature. Values of approx 0 indicate the curve is linear and changing at a steady state. Years in which the second derivative is markedly different from zero are those in which something is happening to alter the rate of population change." (Fewster et al 2000) -->

<!-- ```{r derivative 2, echo = TRUE } -->

<!-- deriv_func <- splinefun(x = fit_obs_df$annual_time_step, y = fit_obs_df$fitted) -->
<!-- deriv_fitted <- as.data.frame(x = deriv_func(seq(1, 290, 1), deriv = 2)) -->

<!-- deriv_df <- cbind(ts = fit_obs_df$annual_time_step,d = deriv_fitted) -->

<!-- ggplot(deriv_df, aes(ts, `deriv_func(seq(1, 290, 1), deriv = 2)`))+ -->
<!--   geom_line() -->


<!-- gratia_derivs_2 <- derivatives(selected_mods[[i]], order = 2,  -->
<!--                     n = 290, interval = "confidence") -->

<!-- draw(gratia_derivs_1) -->
<!-- draw(gratia_derivs_2) -->


<!-- ``` -->

<!-- ```{r misc } -->

<!-- m1 <- gamm(response_var ~ s(annual_time_step)+ pressure_var_f,  -->
<!--           data = pr_gam_inputs, -->
<!--            correlation = corARMA(form = ~ 1|seasonal_var, p = 1)) -->


<!-- intervals(m1$lme, which = "var-cov")$corStruct -->

<!-- AIC(m, m0, m1$lme) -->

<!-- selected_mod <- m1 -->

<!-- test <- data.frame(annual_time_step = pr_gam_inputs$annual_time_step,  -->
<!--                    fitted = selected_mod$gam$fitted.values,  -->
<!--                    observed = pr_gam_inputs$response_var) -->

<!-- ggplot() +  -->
<!--   geom_line(data = test, aes(annual_time_step, fitted)) +  -->
<!--   geom_point(data = test, aes(annual_time_step, observed)) -->


<!-- deriv_func <- splinefun(x = test$annual_time_step, y = test$fitted) -->
<!-- deriv_fitted <- as.data.frame(x = deriv_func(seq(1, 290, 1), deriv = 2)) -->

<!-- deriv_df <- cbind(ts = test$annual_time_step,d = deriv_fitted) -->

<!-- ggplot(deriv_df, aes(ts, `deriv_func(seq(1, 290, 1), deriv = 2)`))+ -->
<!--   geom_line() -->

<!-- # https://petolau.github.io/Analyzing-double-seasonal-time-series-with-GAM-in-R/ -->
<!-- # https://fromthebottomoftheheap.net/2014/05/09/modelling-seasonal-data-with-gam/ -->
<!-- # https://asbates.rbind.io/2019/04/19/time-series-with-spline-based-models/ -->
<!-- # https://asbates.rbind.io/2019/05/03/gams-for-time-series/  -->

<!-- fit <- auto.arima(resid(m1$lme, type = "normalized"), -->
<!--                        stationary = TRUE, seasonal = TRUE) -->

<!-- fit$coef -->

<!-- plot(forecast(fit,h=20)) -->
<!-- summary(fit) -->

<!-- plot(pr_gam_inputs$pressure_var_f, resid(m1), xlab = "disturbance", -->
<!-- ylab = "Residuals") -->

<!-- plot(pr_gam_inputs$pressure_var_c, resid(m1), xlab = "disturbance level", -->
<!-- ylab = "Residuals") -->

<!-- plot(pr_gam_inputs$annual_time_step, resid(m1), xlab = "annual time step", -->
<!-- ylab = "Residuals") -->



<!-- selected_mod <- mmmm -->

<!-- gam.check(selected_mod$gam) -->

<!-- E <- residuals(selected_mod, type = "response") -->
<!-- I1 <- !is.na(pr_gam_inputs$response_var) -->
<!-- Efull <- vector(length = length(pr_gam_inputs$response_var)) -->
<!-- Efull <- NA -->
<!-- Efull[I1] <- E -->
<!-- acf(Efull, na.action = na.pass, -->
<!-- main = "Auto-correlation plot for residuals") -->

<!-- acf(Efull, plot = FALSE) -->
<!-- pacf(Efull, plot = FALSE) -->

<!-- ``` -->


<!-- # Model selection -->

<!-- ```{r model selection } -->

<!-- #Using ANOVA (as per https://fromthebottomoftheheap.net/2014/05/09/modelling-seasonal-data-with-gam/) -->

<!-- # anova(m1, m2$lme, m3, m4$lme, m5, m6$lme) -->

<!-- ## Using AIC -->
<!-- mod_sel_table <- as.data.frame(AIC(m1, m2$lme, m3$lme, m4)) -->
<!-- mod_sel_table -->

<!-- # Define selected model -->

<!-- selected_mod <- m4 -->

<!-- summary(selected_mod) -->

<!-- plot(selected_mod, residuals = TRUE, pch = 19, cex = 0.75) -->

<!-- gam.check(selected_mod) -->
<!-- ``` -->

<!-- ```{r plot selected mods, eval = FALSE} -->

<!-- summary(pr_gam_inputs) -->

<!-- pdat <- data.frame(annual_time_step = c(1:300), -->
<!--                    pressure_var_f = as.factor("disturbance")) -->

<!-- head(pdat) -->

<!-- p <- predict(selected_mod, newdata = pdat, type = "response") -->
<!-- p -->

<!-- pdat <- cbind(pdat,p) -->
<!-- lines(p ~ annual_time_step, data = pdat, col = "pink") -->
<!-- # lines(p1 ~ annual_time_step, data = pdat, col = "red") -->
<!-- # lines(p2 ~ annual_time_step, data = pdat, col = "blue") -->
<!-- legend("topleft", -->
<!--        legend = c("Uncorrelated Errors","AR(1) Errors"), -->
<!--        bty = "n", col = c("red","blue"), lty = 1) -->

<!-- # *** Plot the derivatives and periods of change ---- -->

<!-- m2.d <- Deriv(selected_mod$terms, n = 300) -->
<!-- plot(m2.d, sizer = TRUE, alpha = 0.01) -->

<!-- # Add periods of change to time series -->

<!-- plot(LPI ~ annual_time_step, data = pr_gam_inputs,  -->
<!--      type = "p") -->
<!-- lines(p ~ annual_time_step, data = pdat) -->
<!-- CI <- confint(m2.d, alpha = 0.01) -->
<!-- S <- signifD(p, m2.d$annual_time_step$deriv,  -->
<!--              CI$annual_time_step$upper,  -->
<!--              CI$annual_time_step$lower, -->
<!--              eval = 0) -->
<!-- lines(S$incr ~ annual_time_step, data = pdat, lwd = 3, col = "blue") -->
<!-- lines(S$decr ~ annual_time_step, data = pdat, lwd = 3, col = "red") -->
<!-- # Recreate with ggplot -->

<!-- deriv_plot <- ggplot() + -->
<!--   geom_point(data = gam_inputs, -->
<!--              aes(x = annual_time_step, y = indicator_score), -->
<!--              alpha = 0.3) + -->
<!--   geom_line(aes(x = gam_inputs$annual_time_step, -->
<!--                 y = p2)) + -->
<!--   geom_line(aes(x = gam_inputs$annual_time_step,  -->
<!--                 y = S$incr), col = "blue", size = 2) + -->
<!--   geom_line(aes(x = gam_inputs$annual_time_step,  -->
<!--                 y = S$decr), col = "red", size = 2) + -->
<!--   labs(title = paste(gam_inputs$scenario[1], -->
<!--                      gam_inputs$indicator[1], sep = " ")) -->

<!-- deriv_plot -->

<!-- ggsave(file.path(analysis_plots_folder, -->
<!--                  paste(gam_inputs$indicator[1],"_", gam_inputs$scenario[1],  -->
<!--                        "derivative_plot.png", sep = "")), -->
<!--        deriv_plot,  device = "png") -->

<!-- ``` -->
